---
title: "Cox models - incidence of tumors"
author: "Balazs Koncz"
date: "2024-10-08"
output: html_document
---

Incidence of tumor
I) Any tumor
Predictors: sex, PTV (cont. or cat.), gpca1-10, frequent genes with PTV

#Setup

```{r}
setwd("/media/balazs/WorkL/balazs/Work/UKBiobank/")
library(survival, lib.loc = "/home/balazs/R/x86_64-pc-linux-gnu-library/4.3")
Packages <- c("tidyverse", "data.table", "magrittr", "tidycmprsk", "Rfast", "survminer", "ggsurvfit", "fastmatch", "future.apply", "ComplexHeatmap", "pbapply", "forestmodel", "clusterProfiler", "org.Hs.eg.db", "ggrepel", "ggsci", "scales")
#"fastmatch", "h2o", , , "protr", , , "networkD3", "ggpubr", "RColorBrewer", "Biostrings", , )
lapply(Packages, require, character.only = TRUE)
rm(Packages)
options(dplyr.summarise.inform = FALSE)

pal_npg("nrc")(9)
scales::show_col(pal_npg("nrc")(9))
pal_npg("nrc", alpha = 0.8)(9)
scales::show_col(pal_npg("nrc", alpha = 0.8)(9))
```

#Cox models I.
Predictors: sex, PTV (cont. or cat.), gpca1-10

##Dataset

```{r}
ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

ptvs = grep("PTV", colnames(ukb_data), value = T)
ptvs = ptvs[grepl("MAF", ptvs)]
ptvs = ptvs[!grepl("Genes", ptvs)]

tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]

tumorous = ukb_data %>% 
  filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T) %>% 
  mutate(time = as.numeric((diag_date - birthdate))) %>% 
  mutate(status = 2) %>%
  dplyr::select(eid, sex, cancer_type, time, status, all_of(ptvs), gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10)

nottumorous = ukb_data %>% 
  #dplyr::filter(!eid %in% tumorous$eid) %>% 
  dplyr::filter(cancer_type == "No_cancer") %>%
  distinct(eid, .keep_all = T) %>% 
  mutate(time = case_when(
    death == T ~ as.numeric(date_of_death - birthdate),
    death == F ~ as.numeric(max(date_of_death, na.rm = T) - birthdate)
  )) %>% 
  mutate(status = 1) %>%
  dplyr::select(eid, sex, cancer_type, time, status, all_of(ptvs), gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10)

intersect(tumorous$eid, nottumorous$eid)

ukb_data = rbind(tumorous, nottumorous)

```

##Cox-models

```{r}

i=0
pbsapply(ptvs, function(p) {
  i=i+1
  .GlobalEnv$i = i
  tempdf = ukb_data %>% 
    dplyr::select(eid, sex, cancer_type, time, status, all_of(p), gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10)
  colnames(tempdf)[6] = "PTVvar"
  tempdf %<>% dplyr::select(eid, time, status, PTVvar, sex, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10)
  res.cox <- coxph(Surv(time, status) ~ PTVvar + sex + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = tempdf, id = tempdf$eid)
  fig = forest_model(model = res.cox)
  ggsave(paste0("Plots/001/Cox_1/cox_fp_anytumor_", p, ".jpg"), plot = fig, width = 40, height = 20, units = "cm", dpi = 150)
})

#PTVtsg - more groups
ukb_data$PTVtsg_MAF_g2 = cut(ukb_data$PTVtsg_MAF, breaks = c(0,1,2,3,4,5,max(ukb_data$PTVtsg_MAF)), include.lowest = T, right = F)
Table(ukb_data$PTVtsg_MAF_g2)
levels(ukb_data$PTVtsg_MAF_g2)
res.cox <- coxph(Surv(time, status) ~ PTVtsg_MAF_g2 + sex + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = ukb_data, id = ukb_data$eid)
fig = forest_model(model = res.cox)
fig
ggsave("Plots/001/Cox_1/cox_fp_anytumor_PTVtsg_MAF_5Groups.jpg", plot = fig, width = 40, height = 20, units = "cm", dpi = 150)

```

#Cox models II.
Predictors: sex, PTV (cont. or cat.), gpca1-10, frequent TSG with PTV
##Dataset

```{r}
load("Objects/overrep_ft_01_tsgs_alltumors")

ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

ukb_data$PTVtsg_MAF_g3 = cut(ukb_data$PTVtsg_MAF, breaks = c(0,1,2,max(ukb_data$PTVtsg_MAF)), include.lowest = T, right = F)
Table(ukb_data$PTVtsg_MAF_g3)
ukb_data$PTVtsg_MAF_g4 = cut(ukb_data$PTVtsg_MAF, breaks = c(0,1,2,3,max(ukb_data$PTVtsg_MAF)), include.lowest = T, right = F)
Table(ukb_data$PTVtsg_MAF_g4)

ptvs = grep("PTVtsg_MAF", colnames(ukb_data), value = T)

tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]

tumorous = ukb_data %>% 
  filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T) %>% 
  mutate(time = as.numeric((diag_date - birthdate))) %>% 
  mutate(status = 2) %>%
  dplyr::select(eid, sex, cancer_type, time, status, all_of(ptvs), Genes_PTV_MAF, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10)

nottumorous = ukb_data %>% 
  #dplyr::filter(!eid %in% tumorous$eid) %>% 
  dplyr::filter(cancer_type == "No_cancer") %>%
  distinct(eid, .keep_all = T) %>% 
  mutate(time = case_when(
    death == T ~ as.numeric(date_of_death - birthdate),
    death == F ~ as.numeric(max(date_of_death, na.rm = T) - birthdate)
  )) %>% 
  mutate(status = 1) %>%
  dplyr::select(eid, sex, cancer_type, time, status, all_of(ptvs), Genes_PTV_MAF, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10)

intersect(tumorous$eid, nottumorous$eid)

ukb_data = rbind(tumorous, nottumorous)

mat = pbsapply(overrep_tsgs_ensg, function(genes) {
  grepl(genes, ukb_data$Genes_PTV_MAF)
})
colnames(mat) = overrep_tsgs_hugo
ukb_data = cbind(ukb_data, mat)
rm(mat)
```

##Cox models and forest plots

```{r}
i=0
pbsapply(ptvs, function(p) {
  i=i+1
  .GlobalEnv$i = i
  tempdf = ukb_data %>% dplyr::select(eid, sex, cancer_type, time, status, all_of(p), gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10, all_of(overrep_tsgs_hugo))
  colnames(tempdf)[6] = "PTVvar"
  tempdf %<>% dplyr::select(eid, time, status, PTVvar, sex, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10, all_of(overrep_tsgs_hugo))
  myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(tempdf)[4:ncol(tempdf)], collapse = "+")))
  res.cox <- coxph(myformula, data = tempdf, id = tempdf$eid)
  fig = forest_model(model = res.cox, show_global_p = "bottom", n_logical_true_only = T)
  ggsave(paste0("Plots/001/Cox_2/cox_fp_anytumor_", p, "_freqTSGs.jpg"), plot = fig, width = 20, height = 25, units = "cm", dpi = 150)
})


#Exclude samples with PTVs in BRCA1, CDH1, MLH1, PALB2, MSH2, DNMT3A" "APC, BARD1, BRCA2, ATM, BAP1, TET2, ASXL1, ALX15B, CHEK2
pbsapply(ptvs, function(p) {
  tempdf = ukb_data %>% dplyr::select(eid, sex, cancer_type, time, status, all_of(p), gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10, all_of(overrep_tsgs_hugo))
  colnames(tempdf)[6] = "PTVvar"
  tempdf %<>% dplyr::select(eid, time, status, PTVvar, sex, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10, all_of(overrep_tsgs_hugo))
  tempdf = tempdf[!apply(tempdf[,16:ncol(tempdf)], 1, any),]
  res.cox <- coxph(Surv(time, status) ~ PTVvar + sex + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = tempdf, id = tempdf$eid)
  fig = forest_model(model = res.cox, show_global_p = "bottom")
  ggsave(paste0("Plots/001/Cox_2/cox_fp_anytumor_", p, "_freqTSGs_filtered.jpg"), plot = fig, width = 20, height = 25, units = "cm", dpi = 150)
})

```

##Incidence plot - PTVtsg
Exluding of patients with PTV in selected genes decreases the effect of PTV burden!

```{r}
tempdf = ukb_data[!apply(ukb_data[,21:35], 1, any),]

Table(tempdf$status)
tempdf$inc_event = ifelse(tempdf$status == 1, 0, 1)
tempdf$inc_event = as.factor(tempdf$inc_event)
Table(tempdf$inc_event)
tempdf %<>% mutate(inc_age = round(time/365.25))

tempdf %<>% dplyr::select(eid, sex, inc_age, inc_event, PTVtsg_MAF_g, PTVtsg_MAF_g3, PTVtsg_MAF_g4)


ggcuminc(x = cuminc(Surv(time = inc_age, event = inc_event) ~ PTVtsg_MAF_g + sex, data = tempdf)) +
  add_confidence_interval() +
  scale_ggsurvfit() +
  # annotate(geom = "text", label = paste0("\nfemale: p = ", grayp_f, ", male: p = ", grayp_m), x = 10, y = Inf ) +
  xlab("Time (days)") +
  ggtitle("Cumulative incidence") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"), legend.position = "right") +
  coord_cartesian(xlim = c(40, 90))


#Female
ggcuminc(x = cuminc(Surv(time = inc_age, event = inc_event) ~ PTVtsg_MAF_g, data = tempdf[tempdf$sex == "female",])) +
  add_confidence_interval() +
  scale_ggsurvfit() +
  # annotate(geom = "text", label = paste0("\nfemale: p = ", grayp_f, ", male: p = ", grayp_m), x = 10, y = Inf ) +
  xlab("Time (age)") +
  ggtitle("Cumulative incidence") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"), legend.position = "right") +
  coord_cartesian(xlim = c(40, 90))

#Male
ggcuminc(x = cuminc(Surv(time = inc_age, event = inc_event) ~ PTVtsg_MAF_g, data = tempdf[tempdf$sex == "male",])) +
  add_confidence_interval(type = "ribbon") +
  scale_ggsurvfit() +
  # annotate(geom = "text", label = paste0("\nfemale: p = ", grayp_f, ", male: p = ", grayp_m), x = 10, y = Inf ) +
  xlab("Time (age)") +
  ggtitle("Cumulative incidence") +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"), legend.position = "right") +
  coord_cartesian(xlim = c(40, 90))

```

