---
title: "B_010_Collect_frequent_PTVs_by_tumors_and_genders"
output: html_document
date: "2024-11-12"
---

#Setup

```{r}
setwd("/media/balazs/WorkL/balazs/Work/UKBiobank/")
library(survival, lib.loc = "/home/balazs/R/x86_64-pc-linux-gnu-library/4.3")
Packages <- c("tidyverse", "data.table", "magrittr", "tidycmprsk", "Rfast", "survminer", "ggsurvfit", "fastmatch", "future.apply", "ComplexHeatmap", "pbapply", "forestmodel", "clusterProfiler", "org.Hs.eg.db", "ggrepel", "ggsci", "scales", "ggpubr", "RColorBrewer")
lapply(Packages, require, character.only = TRUE)
rm(Packages)
options(dplyr.summarise.inform = FALSE)
```

#Dataset

```{r}
load("Objects/ukb_data_cancer_final")
geg = fread("Variables/22006.txt") #caucasians
sum(Table(ukb_data_cancer$cancer_type))
ukb_data = ukb_data_cancer %>% 
  filter(eid %fin% as.character(geg$eid)) %>% 
  dplyr::select(eid, birthdate, sex, date_of_death, death, diag_date, cancer_type, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10, gpca11, gpca12, gpca13, gpca14, gpca15, gpca16, gpca17, gpca18, gpca19, gpca20)
rm(ukb_data_cancer, geg)

ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

ptvb_MAF103 = fread("PTVvars/ultra_rare_variants_MAF_all_retained_variants_PTVBurden_final_Shetscores.tsv")
ptvb_MAF103 = ptvb_MAF103[ptvb_MAF103$Patient.ID %fin% ukb_data$eid,]

ptvb_MAF104 = fread("PTVvars/MAF10_4_all_retained_variants_PTVBurden_final_Shetscores.tsv")
ptvb_MAF104 = ptvb_MAF104[ptvb_MAF104$Patient.ID %fin% ukb_data$eid,]

genesFreq = data.frame(ENSEMBL = sort(unique(c(unlist(strsplit(ptvb_MAF103$Genes, ","), use.names = F), unlist(strsplit(ptvb_MAF104$Genes, ","), use.names = F)))))
gene_symbols = bitr(geneID = genesFreq$ENSEMBL, fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
genesFreq$SYMBOL = gene_symbols$SYMBOL[fmatch(genesFreq$ENSEMBL, gene_symbols$ENSEMBL)]
rm(gene_symbols)

ukb_data$PTVgenes103 = ptvb_MAF103$Genes[fmatch(ukb_data$eid, ptvb_MAF103$Patient.ID)]
ukb_data$PTVgenes104 = ptvb_MAF104$Genes[fmatch(ukb_data$eid, ptvb_MAF104$Patient.ID)]
rm(ptvb_MAF103, ptvb_MAF104)

tsg = fread("Objects/tumor_supressor_genes_bioinfo_uth.txt")
genesFreq$tsg = genesFreq$SYMBOL %in% tsg$GeneSymbol
rm(tsg)

save(ukb_data, file = "Objects/010/ukb_data")
```

#PTV gene frequencies
##MAF 10-3
###All UKB

```{r}
ukb_data_unique = ukb_data %>% 
  dplyr::select(eid, PTVgenes103) %>% 
  unique()
genesFreq$n_UKB = nrow(ukb_data_unique)

temp = Table(unlist(lapply(ukb_data_unique$PTVgenes103, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$allUKB = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$allUKB[is.na(genesFreq$allUKB)] = 0
rm(temp)

##Females
ukb_data_female = ukb_data %>% 
  dplyr::filter(sex == "female") %>% 
  dplyr::select(eid, PTVgenes103) %>% 
  unique()
genesFreq$n_UKB_female = nrow(ukb_data_female)

##Males
ukb_data_male = ukb_data %>% 
  dplyr::filter(sex == "male") %>% 
  dplyr::select(eid, PTVgenes103) %>% 
  unique()
genesFreq$n_UKB_male = nrow(ukb_data_male)
rm(ukb_data_female, ukb_data_male, ukb_data_unique)

```

###Not tumorous

```{r}
tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]
v1 = ukb_data %>%
  filter(cancer_type == "No_cancer") %>%
  pull(eid)
v2 = ukb_data %>%
  filter(cancer_type %in% tumors) %>%
  pull(eid)
length(intersect(v1, v2)) #0
rm(v1, v2)

#ALL
nottumorous = ukb_data %>% 
  filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)

temp = Table(unlist(lapply(nottumorous$PTVgenes103, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_notTumorous = nrow(nottumorous)
genesFreq$notTumorous = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$notTumorous[is.na(genesFreq$notTumorous)] = 0
rm(temp, nottumorous)

#FEMALE
nottumorous_female = ukb_data %>% 
  dplyr::filter(sex == "female") %>% 
  dplyr::filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(nottumorous_female$PTVgenes103, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_notTumorous_f = nrow(nottumorous_female)
genesFreq$notTumorous_f = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$notTumorous_f[is.na(genesFreq$notTumorous_f)] = 0
rm(temp, nottumorous_female)

#MALE
nottumorous_male = ukb_data %>% 
  dplyr::filter(sex == "male") %>% 
  dplyr::filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(nottumorous_male$PTVgenes103, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_notTumorous_m = nrow(nottumorous_male)
genesFreq$notTumorous_m = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$notTumorous_m[is.na(genesFreq$notTumorous_m)] = 0
rm(temp, nottumorous_male)

```

###Tumorous
####Any kind of tumors

```{r}
tumorous = ukb_data %>% 
  filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(tumorous$PTVgenes103, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_tumorous = nrow(tumorous)
genesFreq$tumorous = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$tumorous[is.na(genesFreq$tumorous)] = 0
rm(temp, tumorous)

#FEMALE
tumorous_female = ukb_data %>% 
  dplyr::filter(sex == "female") %>%
  dplyr::filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(tumorous_female$PTVgenes103, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_tumorous_f = nrow(tumorous_female)
genesFreq$tumorous_f = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$tumorous_f[is.na(genesFreq$tumorous_f)] = 0
rm(temp, tumorous_female)

#MALE
tumorous_male = ukb_data %>% 
  dplyr::filter(sex == "male") %>%
  dplyr::filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(tumorous_male$PTVgenes103, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_tumorous_m = nrow(tumorous_male)
genesFreq$tumorous_m = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$tumorous_m[is.na(genesFreq$tumorous_m)] = 0
rm(temp, tumorous_male)
```

####Tumors separately

```{r}
tumor_dfs = pblapply(tumors, function(t) {
  ukb_data %>% 
    filter(cancer_type == t) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T)
})
names(tumor_dfs) = tumors

plan(multisession(workers = 8))
tumor_freqmat = future_lapply(tumors, function(t) {
  temp = Table(unlist(lapply(tumor_dfs[[t]]$PTVgenes103, function(g) {
    unique(strsplit(g, ",")[[1]])
  })))
  temp2 = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
  temp2[is.na(temp2)] = 0
  out = cbind(rep(nrow(tumor_dfs[[t]]), length(temp2)), temp2)
  colnames(out) = c(paste0("n_", t), t)
  out
}, future.seed=TRUE)
tumor_freqmat = do.call(cbind, tumor_freqmat)
genesFreq = cbind(genesFreq, tumor_freqmat)
rm(tumor_dfs, tumor_freqmat)


```

####Tumors and genders separately

```{r}
tumor_freq_data = expand.grid(cancer_type = tumors, sex = c("female", "male"), stringsAsFactors = F)
tumor_freq_data$n = apply(tumor_freq_data, 1, function(x) {
  ukb_data %>% 
    filter(cancer_type == x[1], sex == x[2]) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T) %>% 
    nrow()
})
tumor_freq_data %<>% 
  filter(n>=100) %>% 
  mutate(TS = paste(cancer_type, sex, sep = "_"), .before = "n")

tumor_gender_dfs = pbapply(tumor_freq_data, 1, function(x) {
  ukb_data %>% 
    filter(cancer_type == x[1], sex == x[2]) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T)
})
names(tumor_gender_dfs) = tumor_freq_data$TS



plan(multisession(workers = 8))
tumor_gender_freqmat = future_lapply(tumor_freq_data$TS, function(t) {
  temp = Table(unlist(lapply(tumor_gender_dfs[[t]]$PTVgenes103, function(g) {
    unique(strsplit(g, ",")[[1]])
  })))
  temp2 = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
  temp2[is.na(temp2)] = 0
  out = cbind(rep(nrow(tumor_gender_dfs[[t]]), length(temp2)), temp2)
  colnames(out) = c(paste0("n_", t), t)
  out
}, future.seed=TRUE)
tumor_gender_freqmat = do.call(cbind, tumor_gender_freqmat)
genesFreq = cbind(genesFreq, tumor_gender_freqmat)
rm(tumor_gender_dfs, tumor_gender_freqmat)

saveRDS(genesFreq, file = "Res/010/genesFreq_all_103.rds")

```

##MAF 10-4

###All UKB

```{r}
genesFreq %<>% dplyr::select(ENSEMBL, SYMBOL, tsg)
ukb_data_unique = ukb_data %>% 
  dplyr::select(eid, PTVgenes104) %>% 
  unique()
genesFreq$n_UKB = nrow(ukb_data_unique)

temp = Table(unlist(lapply(ukb_data_unique$PTVgenes104, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$allUKB = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$allUKB[is.na(genesFreq$allUKB)] = 0
rm(temp)

##Females
ukb_data_female = ukb_data %>% 
  dplyr::filter(sex == "female") %>% 
  dplyr::select(eid, PTVgenes104) %>% 
  unique()
genesFreq$n_UKB_female = nrow(ukb_data_female)

##Males
ukb_data_male = ukb_data %>% 
  dplyr::filter(sex == "male") %>% 
  dplyr::select(eid, PTVgenes104) %>% 
  unique()
genesFreq$n_UKB_male = nrow(ukb_data_male)
rm(ukb_data_female, ukb_data_male, ukb_data_unique)

```

###Not tumorous

```{r}
tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]
v1 = ukb_data %>%
  filter(cancer_type == "No_cancer") %>%
  pull(eid)
v2 = ukb_data %>%
  filter(cancer_type %in% tumors) %>%
  pull(eid)
length(intersect(v1, v2)) #0
rm(v1, v2)

#ALL
nottumorous = ukb_data %>% 
  filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)

temp = Table(unlist(lapply(nottumorous$PTVgenes104, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_notTumorous = nrow(nottumorous)
genesFreq$notTumorous = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$notTumorous[is.na(genesFreq$notTumorous)] = 0
rm(temp, nottumorous)

#FEMALE
nottumorous_female = ukb_data %>% 
  dplyr::filter(sex == "female") %>% 
  dplyr::filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(nottumorous_female$PTVgenes104, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_notTumorous_f = nrow(nottumorous_female)
genesFreq$notTumorous_f = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$notTumorous_f[is.na(genesFreq$notTumorous_f)] = 0
rm(temp, nottumorous_female)

#MALE
nottumorous_male = ukb_data %>% 
  dplyr::filter(sex == "male") %>% 
  dplyr::filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(nottumorous_male$PTVgenes104, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_notTumorous_m = nrow(nottumorous_male)
genesFreq$notTumorous_m = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$notTumorous_m[is.na(genesFreq$notTumorous_m)] = 0
rm(temp, nottumorous_male)

```

###Tumorous
####Any kind of tumors

```{r}
tumorous = ukb_data %>% 
  filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(tumorous$PTVgenes104, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_tumorous = nrow(tumorous)
genesFreq$tumorous = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$tumorous[is.na(genesFreq$tumorous)] = 0
rm(temp, tumorous)

#FEMALE
tumorous_female = ukb_data %>% 
  dplyr::filter(sex == "female") %>%
  dplyr::filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(tumorous_female$PTVgenes104, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_tumorous_f = nrow(tumorous_female)
genesFreq$tumorous_f = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$tumorous_f[is.na(genesFreq$tumorous_f)] = 0
rm(temp, tumorous_female)

#MALE
tumorous_male = ukb_data %>% 
  dplyr::filter(sex == "male") %>%
  dplyr::filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(tumorous_male$PTVgenes104, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_tumorous_m = nrow(tumorous_male)
genesFreq$tumorous_m = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$tumorous_m[is.na(genesFreq$tumorous_m)] = 0
rm(temp, tumorous_male)
```

####Tumors separately

```{r}
tumor_dfs = pblapply(tumors, function(t) {
  ukb_data %>% 
    filter(cancer_type == t) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T)
})
names(tumor_dfs) = tumors

plan(multisession(workers = 8))
tumor_freqmat = future_lapply(tumors, function(t) {
  temp = Table(unlist(lapply(tumor_dfs[[t]]$PTVgenes104, function(g) {
    unique(strsplit(g, ",")[[1]])
  })))
  temp2 = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
  temp2[is.na(temp2)] = 0
  out = cbind(rep(nrow(tumor_dfs[[t]]), length(temp2)), temp2)
  colnames(out) = c(paste0("n_", t), t)
  out
}, future.seed=TRUE)
tumor_freqmat = do.call(cbind, tumor_freqmat)
genesFreq = cbind(genesFreq, tumor_freqmat)
rm(tumor_dfs, tumor_freqmat)


```

####Tumors and genders separately

```{r}
tumor_freq_data = expand.grid(cancer_type = tumors, sex = c("female", "male"), stringsAsFactors = F)
tumor_freq_data$n = apply(tumor_freq_data, 1, function(x) {
  ukb_data %>% 
    filter(cancer_type == x[1], sex == x[2]) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T) %>% 
    nrow()
})
tumor_freq_data %<>% 
  filter(n>=100) %>% 
  mutate(TS = paste(cancer_type, sex, sep = "_"), .before = "n")

tumor_gender_dfs = pbapply(tumor_freq_data, 1, function(x) {
  ukb_data %>% 
    filter(cancer_type == x[1], sex == x[2]) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T)
})
names(tumor_gender_dfs) = tumor_freq_data$TS



plan(multisession(workers = 8))
tumor_gender_freqmat = future_lapply(tumor_freq_data$TS, function(t) {
  temp = Table(unlist(lapply(tumor_gender_dfs[[t]]$PTVgenes104, function(g) {
    unique(strsplit(g, ",")[[1]])
  })))
  temp2 = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
  temp2[is.na(temp2)] = 0
  out = cbind(rep(nrow(tumor_gender_dfs[[t]]), length(temp2)), temp2)
  colnames(out) = c(paste0("n_", t), t)
  out
}, future.seed=TRUE)
tumor_gender_freqmat = do.call(cbind, tumor_gender_freqmat)
genesFreq = cbind(genesFreq, tumor_gender_freqmat)
rm(tumor_gender_dfs, tumor_gender_freqmat)

saveRDS(genesFreq, file = "Res/010/genesFreq_all_104.rds")
genesFreq = readRDS("Res/010/genesFreq_all_104.rds")
```

#Overrepresented genes in tumors
##MAF 10-3
G6PC1 (ENSG00000131482), Rectum Male, OR: 5.370108 - evidences, literature?
ASXL1 (Leukemia, Leukemia_male), 
BRCA2 (Breast_female), 
PALB2 (Breast_female), 
ASXL1, BRCA2, PALB2 (tumorous)

```{r}
genesFreq = readRDS("Res/010/genesFreq_all_103.rds")
cats = c("tumorous", colnames(genesFreq)[seq(21,69,2)])
n_nottumorous = genesFreq$n_notTumorous[1]
genesFreq12 = genesFreq %>% dplyr::select(ENSEMBL, SYMBOL)

plan(multisession(workers = 12))
res_fish_1 = future_lapply(cats, function(x) {
  ind1 = fmatch(x, colnames(genesFreq))
  ind2 = fmatch(paste0("n_", x), colnames(genesFreq))
  ftpOR = t(apply(genesFreq, 1, function(y) {
    ft = fisher.test(rbind(c(as.numeric(y[ind1]), as.numeric(y[9])), c(as.numeric(y[ind2]) - as.numeric(y[ind1]), n_nottumorous - as.numeric(y[9]))))
    c(unname(ft$estimate), ft$p.value)
  }))
  ftpOR %<>% as.data.frame() %>% set_colnames(c("OR", "P")) %>% mutate(category = x, .before = "OR") %>% mutate(Padj = p.adjust(p = P, method = "BH"))
  bind_cols(genesFreq12, ftpOR)
})
res_fish_1 = bind_rows(res_fish_1)

#####
cats2 = colnames(genesFreq)[seq(71,ncol(genesFreq),2)]
res_fish_2 = future_lapply(cats2, function(x) {
  ind1 = fmatch(x, colnames(genesFreq))
  ind2 = fmatch(paste0("n_", x), colnames(genesFreq))
  if(grepl("_female", x)) {
    ftpOR = t(apply(genesFreq, 1, function(y) {
      ft = fisher.test(rbind(c(as.numeric(y[ind1]), as.numeric(y[11])), c(as.numeric(y[ind2]) - as.numeric(y[ind1]), as.numeric(y[10]) - as.numeric(y[11]))))
      c(unname(ft$estimate), ft$p.value)
    }))
  } else {
    ftpOR = t(apply(genesFreq, 1, function(y) {
      ft = fisher.test(rbind(c(as.numeric(y[ind1]), as.numeric(y[13])), c(as.numeric(y[ind2]) - as.numeric(y[ind1]), as.numeric(y[12]) - as.numeric(y[13]))))
      c(unname(ft$estimate), ft$p.value)
    }))
  }
  ftpOR %<>% as.data.frame() %>% set_colnames(c("OR", "P")) %>% mutate(category = x, .before = "OR") %>% mutate(Padj = p.adjust(p = P, method = "BH"))
  bind_cols(genesFreq12, ftpOR)
})
res_fish_2 = bind_rows(res_fish_2)

res_fishtest = rbind(res_fish_1, res_fish_2)

save(res_fishtest, file = "Res/010/res_fishtest_103")


View(genesFreq[genesFreq$ENSEMBL == "ENSG00000104804",])
```


##MAF 10-4

```{r}
genesFreq = readRDS("Res/010/genesFreq_all_104.rds")
cats = c("tumorous", colnames(genesFreq)[seq(21,69,2)])
n_nottumorous = genesFreq$n_notTumorous[1]
genesFreq12 = genesFreq %>% dplyr::select(ENSEMBL, SYMBOL)

plan(multisession(workers = 12))
res_fish_1 = future_lapply(cats, function(x) {
  ind1 = fmatch(x, colnames(genesFreq))
  ind2 = fmatch(paste0("n_", x), colnames(genesFreq))
  ftpOR = t(apply(genesFreq, 1, function(y) {
    mtx = rbind(c(as.numeric(y[ind1]), as.numeric(y[ind2]) - as.numeric(y[ind1])), c(as.numeric(y[9]), n_nottumorous - as.numeric(y[9])))
    ft = fisher.test(mtx)
    c(mtx, unname(ft$estimate), ft$p.value)
  }))
  ftpOR %<>% as.data.frame() %>% set_colnames(c("caseY", "controlY", "caseN", "controlN", "OR", "P")) %>% mutate(category = x, .before = "caseY") %>% mutate(Padj = p.adjust(p = P, method = "fdr"))
  bind_cols(genesFreq12, ftpOR)
})
res_fish_1 = bind_rows(res_fish_1)
rm(cats, n_nottumorous)

#####
cats2 = colnames(genesFreq)[c(17,19,seq(71,ncol(genesFreq),2))]
n_nottumorous_f = genesFreq$n_notTumorous_f[1]
n_nottumorous_m = genesFreq$n_notTumorous_m[1]
res_fish_2 = future_lapply(cats2, function(x) {
  ind1 = fmatch(x, colnames(genesFreq))
  ind2 = fmatch(paste0("n_", x), colnames(genesFreq))
  if(grepl("_f", x) | grepl("_female", x)) {
    ftpOR = t(apply(genesFreq, 1, function(y) {
      mtx_f = rbind(c(as.numeric(y[ind1]), as.numeric(y[ind2]) - as.numeric(y[ind1])), c(as.numeric(y[11]), n_nottumorous_f - as.numeric(y[11])))
      ft = fisher.test(mtx_f)
      c(mtx_f, unname(ft$estimate), ft$p.value)
    }))
  } else {
    ftpOR = t(apply(genesFreq, 1, function(y) {
      mtx_m = rbind(c(as.numeric(y[ind1]), as.numeric(y[ind2]) - as.numeric(y[ind1])), c(as.numeric(y[13]), n_nottumorous_m - as.numeric(y[13])))
      ft = fisher.test(mtx_m)
      c(mtx_m, unname(ft$estimate), ft$p.value)
    }))
  }
  ftpOR %<>% as.data.frame() %>% set_colnames(c("caseY", "controlY", "caseN", "controlN", "OR", "P")) %>% mutate(category = x, .before = "caseY") %>% mutate(Padj = p.adjust(p = P, method = "fdr"))
  bind_cols(genesFreq12, ftpOR)
})
res_fish_2 = bind_rows(res_fish_2)

rm(cats2, n_nottumorous_f, n_nottumorous_m)

res_fishtest = rbind(res_fish_1, res_fish_2)

saveRDS(res_fishtest, file = "Res/010/res_fishtest_104.rds")

res_fishtest = readRDS("Res/010/res_fishtest_104.rds")
View(genesFreq[genesFreq$ENSEMBL == "ENSG00000104804",])
```

#GO enrichment analysis

```{r}
res_fishtest$ENTREZID = mapIds(org.Hs.eg.db, keys = res_fishtest$ENSEMBL, keytype="ENSEMBL", column = "ENTREZID")
genesFreq12$ENTREZID = mapIds(org.Hs.eg.db, keys = genesFreq12$ENSEMBL, keytype="ENSEMBL", column = "ENTREZID")
res_fishtest %>% filter(Padj < 0.2) %>% pull(category) %>% table() %>% sort(decreasing = T)
t = "tumorous"
t = "Leukemia"
t = "Breast"
t = "Colon_male"
t = "Prostate"
t = "Lung"

ggo <- groupGO(gene = res_fishtest$ENTREZID[res_fishtest$OR > 1 & res_fishtest$Padj < .2 & res_fishtest$category == t],
               OrgDb = org.Hs.eg.db,
               ont = "CC",
               level = 3,
               readable = TRUE)
head(ggo)
ggo@result %>% View()


#pvalueCutoff = 0.01
#qvalueCutoff = 0.05
ego <- enrichGO(gene = res_fishtest$ENTREZID[res_fishtest$OR > 1 & res_fishtest$Padj < .2 & res_fishtest$category == t],
                universe = genesFreq12$ENTREZID,
                OrgDb = org.Hs.eg.db,
                ont = "CC",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.1,
                qvalueCutoff = 0.1,
                readable = TRUE)
goplot(ego)
ego@result
head(ego)


```

#Melanoma
Melanoma_female: some irrelevant terms (BP: sphingomyelin metabolic process, response to dsRNA, response to exogenous dsRNA, MF: phosphatidylinositol bisphosphate binding, protein sequestering activity, interleukin-1 receptor binding, molecular sequestering activity, phosphatidylinositol phosphate binding, microtubule plus-end binding, ligase activity)
Melanoma_male: -

```{r}
genesFreq = readRDS("Res/010/genesFreq_all_103.rds")
res_fishtest = readRDS("Res/010/res_fishtest_104.rds")
res_fishtest$ENTREZID = mapIds(org.Hs.eg.db, keys = res_fishtest$ENSEMBL, keytype="ENSEMBL", column = "ENTREZID")
genesFreq12$ENTREZID = mapIds(org.Hs.eg.db, keys = genesFreq12$ENSEMBL, keytype="ENSEMBL", column = "ENTREZID")

ego <- enrichGO(gene = res_fishtest %>% dplyr::filter(OR > 1, P < 0.2, category == "Melanoma_female") %>% pull(ENTREZID) %>% na.omit(),
                universe = genesFreq$ENTREZID,
                OrgDb = org.Hs.eg.db,
                ont = "ALL",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.2,
                qvalueCutoff = 0.2,
                readable = TRUE)
goplot(ego)
ego@result %>% View()
head(ego)

```

##Incidence

#Dataset

```{r}
res_fishtest = readRDS("Res/010/res_fishtest_104.rds")
ptvb_MAF104 = fread("PTVvars/MAF10_4_all_retained_variants_PTVBurden_final_Shetscores.tsv")
ptvb_MAF104 %<>% dplyr::select(Patient.ID, Genes) %>% set_colnames(c("eid", "ptvgenes"))
broca_crp = c("ALK", "APC", "ATM", "ATR", "AXIN2", "BAP1", "BARD1", "BMPR1A", "BRCA1", "BRCA2", "BRIP1", "CDH1", "CDK4", "CDK12", "CDKN2A", "CHEK2", "CTNNA1", "DICER1", "EPCAM", "FANCM", "FH", "FLCN", "GEN1", "GREM1", "HOXB13", "MEN1", "MET", "MITF", "MLH1", "MLH3", "MSH2", "MSH3", "MSH6", "MUTYH", "NBN", "NF1", "NF2", "NTHL1", "PALB2", "PHOX2B", "PIK3CA", "PMS2", "POLD1", "POLE", "POT1", "PRKAR1A", "PTCH1", "PTEN", "RAD51B", "RAD51C", "RAD51D", "RB1", "RECQL", "RET", "RNF43", "RPS20", "SDHB", "SDHC", "SDHD", "SMAD4", "SMARCA4", "TP53", "TSC1", "TSC2", "VHL")

tempensgs = res_fishtest %>% filter(!SYMBOL %in% broca_crp) %>% filter(category == "Melanoma", OR > 1, P < 0.05) %>% pull(ENSEMBL) #453
ptvb_MAF104$PTVb = sapply(ptvb_MAF104$ptvgenes, function(x) sum(unique(unlist(strsplit(x, ","), use.names = F)) %fin% tempensgs))

ukb_data_raw = readRDS("Objects/ukb_data.rds")
geg = fread("Variables/22006.txt") #caucasians
ukb_dataF = ukb_data_raw %>% 
  filter(eid %fin% as.character(geg$eid))
rm(geg)

ukb_dataF$sex = factor(ukb_dataF$sex, levels = c(0,1), labels = c("female", "male"))
sel_gpcas = paste0("gpca", 1:20)
ukb_dataF %<>% dplyr::select(eid, birthdate, sex, date_of_death, death, diag_date, cancer_type, all_of(sel_gpcas))
ukb = ukb_dataF %>% left_join(ptvb_MAF104, by = "eid") %>% dplyr::select(-ptvgenes)

melanomaDF = ukb %>% 
  filter(cancer_type == "Melanoma") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T) %>% 
  mutate(time = as.numeric((diag_date - birthdate))) %>% 
  mutate(status = 2) %>%
  dplyr::select(eid, sex, cancer_type, time, status, PTVb, all_of(sel_gpcas))
nottumorous = ukb %>% 
  dplyr::filter(cancer_type == "No_cancer") %>%
  distinct(eid, .keep_all = T) %>% 
  mutate(time = case_when(
    death == T ~ as.numeric(date_of_death - birthdate),
    death == F ~ as.numeric(max(date_of_death, na.rm = T) - birthdate)
  )) %>% 
  mutate(status = 1) %>%
  dplyr::select(eid, sex, cancer_type, time, status, PTVb, all_of(sel_gpcas))

ukb = rbind(melanomaDF, nottumorous)

ukb$PTVG = ifelse(ukb$PTVb == 0, 0, 1)
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(c("sex", "PTVG", sel_gpcas), collapse = "+")))
res.cox <- coxph(myformula, data = ukb, id = ukb$eid)
summary(res.cox)

forest_model(model = res.cox)
table(ukb$PTVb, ukb$cancer_type)
```

##Melanoma female, male

```{r}
res_fishtest = readRDS("Res/010/res_fishtest_104.rds")
ptvb_MAF104 = fread("PTVvars/MAF10_4_all_retained_variants_PTVBurden_final_Shetscores.tsv")
ptvb_MAF104 %<>% dplyr::select(Patient.ID, Genes) %>% set_colnames(c("eid", "ptvgenes"))
broca_crp = c("ALK", "APC", "ATM", "ATR", "AXIN2", "BAP1", "BARD1", "BMPR1A", "BRCA1", "BRCA2", "BRIP1", "CDH1", "CDK4", "CDK12", "CDKN2A", "CHEK2", "CTNNA1", "DICER1", "EPCAM", "FANCM", "FH", "FLCN", "GEN1", "GREM1", "HOXB13", "MEN1", "MET", "MITF", "MLH1", "MLH3", "MSH2", "MSH3", "MSH6", "MUTYH", "NBN", "NF1", "NF2", "NTHL1", "PALB2", "PHOX2B", "PIK3CA", "PMS2", "POLD1", "POLE", "POT1", "PRKAR1A", "PTCH1", "PTEN", "RAD51B", "RAD51C", "RAD51D", "RB1", "RECQL", "RET", "RNF43", "RPS20", "SDHB", "SDHC", "SDHD", "SMAD4", "SMARCA4", "TP53", "TSC1", "TSC2", "VHL")

#Female
tempensgs = res_fishtest %>% filter(!SYMBOL %in% broca_crp) %>% filter(category == "Melanoma_female", OR > 1, P < 0.05) %>% pull(ENSEMBL) #407 (406)
#Male
tempensgs = res_fishtest%>% filter(!SYMBOL %in% broca_crp) %>% filter(category == "Melanoma_male", OR > 1, P < 0.05) %>% pull(ENSEMBL) #402 (398)
ptvb_MAF104$PTVb = sapply(ptvb_MAF104$ptvgenes, function(x) sum(unique(unlist(strsplit(x, ","), use.names = F)) %fin% tempensgs))

tempdfU = readRDS("/media/balazs/WorkL/balazs/Work/UKBiobank/Objects/014/dfs/Melanoma_female.rds")
tempdfU = readRDS("/media/balazs/WorkL/balazs/Work/UKBiobank/Objects/014/dfs/Melanoma_male.rds")

sel_gpcas = paste0("gpca", 1:20)
ukb = tempdfU %>% left_join(ptvb_MAF104, by = "eid") %>% dplyr::select(-ptvgenes)
ukb$PTVG = ifelse(ukb$PTVb == 0, 0, 1)
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(c("PTVG", sel_gpcas), collapse = "+")))
res.cox <- coxph(myformula, data = ukb, id = ukb$eid)
forest_model(model = res.cox)

table(ukb$PTVb, ukb$cancer_type)
```



```{r}

tempdfU = readRDS("/media/balazs/WorkL/balazs/Work/UKBiobank/Objects/014/dfs/Melanoma.rds")
ukb = tempdfU %>% left_join(ptv_dataDF, by = "eid")
ukb$PTVG = ifelse(ukb$PTV == 0, 0, 1)
if(length(unique(ukb$PTVG)) > 1) {ukb$PTVG = as.factor(ukb$PTVG)}
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(c("PTVG", sel_gpcas), collapse = "+")))
res.cox <- coxph(myformula, data = ukb, id = ukb$eid)
```

