---
title: "008_Melanoma_PTV_AP"
output: html_document
date: "2024-11-08"
---

#Setup

```{r}
setwd("/media/balazs/WorkL/balazs/Work/UKBiobank/")
Packages <- c("tidyverse", "data.table", "magrittr", "tidycmprsk", "Rfast", "survminer", "ggsurvfit", "fastmatch", "future.apply", "ComplexHeatmap", "pbapply", "forestmodel", "clusterProfiler", "org.Hs.eg.db", "ggrepel", "ggsci", "scales", "ggpubr", "RColorBrewer", "survival")
lapply(Packages, require, character.only = TRUE)
rm(Packages)
options(dplyr.summarise.inform = FALSE)

```

#Functions generate Cox model and Forest plot

```{r}
myFunForestPlot = function(dataset, geneset, tumor, gender, gpca_nb = 10, threads) {
  #plan(future::cluster, workers = threads)
  dataset$PTVburden = sapply(dataset$PTVgenes, function(x) {
    g = strsplit(x, ",")[[1]]
    sum(g %fin% geneset)
  })
  tumorous = dataset %>% 
    dplyr::filter(cancer_type %in% tumor) %>% 
    dplyr::filter(sex %in% gender) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T) %>% 
    mutate(time = as.numeric((diag_date - birthdate))) %>% 
    mutate(status = 2)
  nottumorous = dataset %>% 
    dplyr::filter(cancer_type == "No_cancer") %>%
    dplyr::filter(sex %in% gender) %>%
    distinct(eid, .keep_all = T) %>% 
    mutate(time = case_when(
      death == T ~ as.numeric(date_of_death - birthdate),
      death == F ~ as.numeric(max(date_of_death, na.rm = T) - birthdate)
    )) %>% 
    mutate(status = 1)
  dataset2 = rbind(tumorous, nottumorous)
  dataset2$PTVburdenG = as.factor(ifelse(dataset2$PTVburden == 0, 0, 1))
  #Select predictors
  sel_gpcas = paste0("gpca", 1:gpca_nb)
  if(table(dataset2$PTVburdenG[dataset2$cancer_type != "No_cancer"])["1"] != 0 & length(gender) == 2) {
    dataset2 %<>% dplyr::select(eid, time, status, PTVburdenG, sex, all_of(sel_gpcas))
  } else if(table(dataset2$PTVburdenG[dataset2$cancer_type != "No_cancer"])["1"] != 0 & length(gender) == 1) {
    dataset2 %<>% dplyr::select(eid, time, status, PTVburdenG, all_of(sel_gpcas))
  } else if(table(dataset2$PTVburdenG[dataset2$cancer_type != "No_cancer"])["1"] == 0 & length(gender) == 2) {
    dataset2 %<>% dplyr::select(eid, time, status, sex, all_of(sel_gpcas))
  } else {
    dataset2 %<>% dplyr::select(eid, time, status, all_of(sel_gpcas))
  }
  myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(dataset2)[4:ncol(dataset2)], collapse = "+")))
  res.cox <- coxph(myformula, data = dataset2, id = dataset2$eid)
  fm = forest_model(model = res.cox, return_data = T)
  if(length(gender) == 2) {
    tt = paste0(tumor, ", BOTH")
  } else {
    tt = paste0(tumor, ", ", gender)
  }
  fm$plot + labs(title = tt)
}

myFunCoxStats = function(dataset, geneset, tumor, gender, gpca_nb = 10, threads) {
  #plan(future::cluster, workers = threads)
  #plan(multisession(workers = threads))
  dataset$PTVburden = sapply(dataset$PTVgenes, function(x) {
    g = strsplit(x, ",")[[1]]
    sum(g %fin% geneset)
  })
  tumorous = dataset %>% 
    dplyr::filter(cancer_type %in% tumor) %>% 
    dplyr::filter(sex %in% gender) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T) %>% 
    mutate(time = as.numeric((diag_date - birthdate))) %>% 
    mutate(status = 2)
  nottumorous = dataset %>% 
    dplyr::filter(cancer_type == "No_cancer") %>%
    dplyr::filter(sex %in% gender) %>%
    distinct(eid, .keep_all = T) %>% 
    mutate(time = case_when(
      death == T ~ as.numeric(date_of_death - birthdate),
      death == F ~ as.numeric(max(date_of_death, na.rm = T) - birthdate)
    )) %>% 
    mutate(status = 1)
  dataset2 = rbind(tumorous, nottumorous)
  dataset2$PTVburdenG = as.factor(ifelse(dataset2$PTVburden == 0, 0, 1))
  #Select predictors
  sel_gpcas = paste0("gpca", 1:gpca_nb)
  if(table(dataset2$PTVburdenG[dataset2$cancer_type != "No_cancer"])["1"] != 0 & length(gender) == 2) {
    dataset2 %<>% dplyr::select(eid, time, status, PTVburdenG, sex, all_of(sel_gpcas))
  } else if(table(dataset2$PTVburdenG[dataset2$cancer_type != "No_cancer"])["1"] != 0 & length(gender) == 1) {
    dataset2 %<>% dplyr::select(eid, time, status, PTVburdenG, all_of(sel_gpcas))
  } else if(table(dataset2$PTVburdenG[dataset2$cancer_type != "No_cancer"])["1"] == 0 & length(gender) == 2) {
    dataset2 %<>% dplyr::select(eid, time, status, sex, all_of(sel_gpcas))
  } else {
    dataset2 %<>% dplyr::select(eid, time, status, all_of(sel_gpcas))
  }
  myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(dataset2)[4:ncol(dataset2)], collapse = "+")))
  res.cox <- coxph(myformula, data = dataset2, id = dataset2$eid)
  coefs = summary(res.cox)$coefficients[,"exp(coef)"]
  names(coefs) = paste0("hr_", names(coefs))
  pvalues = summary(res.cox)$coefficients[,"Pr(>|z|)"]
  names(pvalues) = paste0("p_", names(pvalues))
  out = c(coefs, pvalues)
  return(out)
}

myFunGroupNb = function(dataset, geneset, tumor, gender) {
  dataset$PTVburden = sapply(dataset$PTVgenes, function(x) {
    g = strsplit(x, ",")[[1]]
    sum(g %fin% geneset)
  })
  tumorous = dataset %>% 
    dplyr::filter(cancer_type %in% tumor) %>% 
    dplyr::filter(sex %in% gender) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T) %>% 
    mutate(time = as.numeric((diag_date - birthdate))) %>% 
    mutate(status = 2)
  nottumorous = dataset %>% 
    dplyr::filter(cancer_type == "No_cancer") %>%
    dplyr::filter(sex %in% gender) %>%
    distinct(eid, .keep_all = T) %>% 
    mutate(time = case_when(
      death == T ~ as.numeric(date_of_death - birthdate),
      death == F ~ as.numeric(max(date_of_death, na.rm = T) - birthdate)
    )) %>% 
    mutate(status = 1)
  dataset2 = rbind(tumorous, nottumorous)
  dataset2$PTVburdenG = as.factor(ifelse(dataset2$PTVburden == 0, 0, 1))
  table(dataset2$PTVburdenG, dataset2$cancer_type)
}

```

#Dataset

```{r}
ukb_data_raw = readRDS("Objects/ukb_data.rds")
geg = fread("Variables/22006.txt") #caucasians
ukb_dataF = ukb_data_raw %>% filter(eid %fin% as.character(geg$eid)) #425,654
rm(geg)

ukb_dataF$sex = factor(ukb_dataF$sex, levels = c(0,1), labels = c("female", "male"))
ukb_dataF %<>% filter(s.color != "Black") #425,530
ukb_dataF$s.color[ukb_dataF$s.color == "Do not know"] = NA
ukb_dataF$s.color = factor(ukb_dataF$s.color, levels = c("Fair", "Dark olive", "Light olive", "Brown", "Very fair"))
sel_gpcas = paste0("gpca", 1:20)

ukb_dataF %<>% dplyr::select(eid, birthdate, sex, date_of_death, death, diag_date, c.occurences, cancer_type, s.color, all_of(sel_gpcas))
ukb_dataF %<>% filter_at(.vars = vars(starts_with("gpca")), .vars_predicate = any_vars(!is.na(.)))

nottumorous_all = ukb_dataF %>% 
  dplyr::filter(cancer_type == "No_cancer") %>%
  distinct(eid, .keep_all = T) %>% 
  #filter(is.na(c.occurences)) %>% #New filtering step!!!
  mutate(time = case_when(
    death == T ~ as.numeric(date_of_death - birthdate),
    death == F ~ as.numeric(max(date_of_death, na.rm = T) - birthdate)
  )) %>% 
  mutate(status = 1)

melanoma = ukb_dataF %>% 
  dplyr::filter(cancer_type == "Melanoma") %>% 
  distinct(eid, .keep_all = T) %>%
  mutate(time = as.numeric((diag_date - birthdate))) %>% 
  mutate(status = 2)

rm(ukb_data_raw, ukb_dataF)
ukb_data = rbind(melanoma, nottumorous_all)
```

#MAF 10-4 PTV genes

```{r}
ptvb_MAF104 = fread("PTVvars/MAF10_4_all_retained_variants_PTVBurden_final_Shetscores.tsv")
ukb_data$PTVgenes = ptvb_MAF104$Genes[fmatch(ukb_data$eid, ptvb_MAF104$Patient.ID)]

genes_all = data.frame(ENSEMBL = sort(unique(unlist(strsplit(ptvb_MAF104$Genes, ","), use.names = F))))
gene_symbols = bitr(geneID = genes_all$ENSEMBL, fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
genes_all$SYMBOL = gene_symbols$SYMBOL[fmatch(genes_all$ENSEMBL, gene_symbols$ENSEMBL)]
rm(gene_symbols, ptvb_MAF104)
```

#Gene set 1.
##Antigen processing and presentation

```{r}
load("Objects/go_desc_list")
hugos_app1 = go_desc_list[["antigen processing and presentation"]]
ensgs_app1 = genes_all$ENSEMBL[fmatch(hugos_app1, genes_all$SYMBOL)]
ensgs_app1 = ensgs_app1[!is.na(ensgs_app1)]

mat_app1 = pbsapply(ensgs_app1, function(genes) {
  grepl(genes, ukb_data$PTVgenes)
})
colSums(mat_app1)
ukb_data = cbind(ukb_data, mat_app1)

ukb_data$PTVburden_app1 = apply(ukb_data[,ensgs_app1], 1, sum)
ukb_data$PTVburden_app1G = as.factor(ifelse(ukb_data$PTVburden_app1 == 0, 0, 1))

```

##Cox model

```{r}
tempdf = ukb_data %>% dplyr::select(eid, time, status, PTVburden_app1G, sex, s.color, all_of(sel_gpcas))
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(tempdf)[4:ncol(tempdf)], collapse = "+")))
res.cox <- coxph(myformula, data = tempdf, id = tempdf$eid)
forest_model(model = res.cox)
table(tempdf$status, tempdf$PTVburden_app1G, tempdf$sex)
```

##Which genes
ENSG00000128815, WDFY4, WDFY Family Member 4 (12 patients, 5 f, 7 m): no assoc.
ENSG00000103811, CTSH, Cathepsin H (7 patients, 3 f, 4 m): Increased expression of this gene has been correlated with malignant progression of prostate tumors.
ENSG00000177628, GBA1, Glucosylceramidase Beta 1 (5 patients, 3 f, 2 m): no assoc.
ENSG00000118508, RAB32, RAB32, Member RAS Oncogene Family (3 patients, 2 f, 1 m): Breast Lipid-Rich Carcinoma
ENSG00000109113, RAB34, RAB34, Member RAS Oncogene Family (3 patients, 2 f, 1 m): no assoc.
ENSG00000104938, CLEC4M C-Type Lectin Domain Family 4 Member M (2 patients, 2 f, 0 m): Lung cancer
ENSG00000168118, RAB4A RAB4A, Member RAS Oncogene Family (2 patients, 2 f, 0 m): Chronic Myelomonocytic Leukemia
ENSG00000153563, CD8A, CD8 Subunit Alpha (2 patients, 1 f, 1 m): no assoc.

```{r}
ukb_data_MEL = ukb_data %>% filter(status == 2)
View(cbind(geneID = genes_all$SYMBOL[fmatch(ensgs_app1, genes_all$ENSEMBL)], 
           female = colSums(ukb_data_MEL[ukb_data_MEL$sex == "female",ensgs_app1]), 
           male = colSums(ukb_data_MEL[ukb_data_MEL$sex == "male",ensgs_app1])))

```

#Filtered geneset

```{r}
ensgs_app2 = ensgs_app1[!ensgs_app1 %in% c("ENSG00000166128", "ENSG00000084733", "ENSG00000111540", "ENSG00000169213", "ENSG00000182197", "ENSG00000152932")]
ukb_data$PTVburden_app2 = apply(ukb_data[,ensgs_app2], 1, sum)
ukb_data$PTVburden_app2G = as.factor(ifelse(ukb_data$PTVburden_app2 == 0, 0, 1))

##Cox model
tempdf = ukb_data %>% dplyr::select(eid, time, status, PTVburden_app2G, sex, s.color, all_of(sel_gpcas))
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(tempdf)[4:ncol(tempdf)], collapse = "+")))
res.cox <- coxph(myformula, data = tempdf, id = tempdf$eid)
forest_model(model = res.cox)
table(tempdf$status, tempdf$PTVburden_app2G, tempdf$sex)

```

#Without BROCA patients

```{r}
broca_crp = c("ALK", "APC", "ATM", "ATR", "AXIN2", "BAP1", "BARD1", "BMPR1A", "BRCA1", "BRCA2", "BRIP1", "CDH1", "CDK4", "CDK12", "CDKN2A", "CHEK2", "CTNNA1", "DICER1", "EPCAM", "FANCM", "FH", "FLCN", "GEN1", "GREM1", "HOXB13", "MEN1", "MET", "MITF", "MLH1", "MLH3", "MSH2", "MSH3", "MSH6", "MUTYH", "NBN", "NF1", "NF2", "NTHL1", "PALB2", "PHOX2B", "PIK3CA", "PMS2", "POLD1", "POLE", "POT1", "PRKAR1A", "PTCH1", "PTEN", "RAD51B", "RAD51C", "RAD51D", "RB1", "RECQL", "RET", "RNF43", "RPS20", "SDHB", "SDHC", "SDHD", "SMAD4", "SMARCA4", "TP53", "TSC1", "TSC2", "VHL")

ensgs_broca = genes_all$ENSEMBL[fmatch(broca_crp, genes_all$SYMBOL)]
ensgs_broca = ensgs_broca[!is.na(ensgs_broca)]
mat_broca = pbsapply(ensgs_broca, function(genes) {
  grepl(genes, ukb_data$PTVgenes)
})
keep_ind = apply(mat_broca, 1, function(z) !any(z))

ukb_dataF = ukb_data[keep_ind,]
tempdf = ukb_dataF %>% dplyr::select(eid, time, status, PTVburden_app1G, sex, s.color, all_of(sel_gpcas))
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(tempdf)[4:ncol(tempdf)], collapse = "+")))
res.cox <- coxph(myformula, data = tempdf, id = tempdf$eid)
forest_model(model = res.cox)
table(tempdf$status, tempdf$PTVburden_app1G, tempdf$sex) #1 man is excluded
```

#Other gene sets

```{r}
load("Objects/go_desc_list")
ensgs_all_app = grep("antigen processing", names(go_desc_list), value = T)
go_desc_list = go_desc_list[ensgs_all_app]
hugos = unique(unlist(go_desc_list, use.names = F))
ensgs = genes_all$ENSEMBL[fmatch(hugos, genes_all$SYMBOL)]
ensgs = ensgs[!is.na(ensgs)] #108 genes

mat_app2 = pbsapply(ensgs, function(genes) {
  grepl(genes, ukb_data$PTVgenes)
})
colSums(mat_app2)
ukb_data = cbind(ukb_data, mat_app2)

#All genes - 108 genes
ukb_data$PTVburden_app3 = apply(ukb_data[,ensgs], 1, sum)
ukb_data$PTVburden_app3G = as.factor(ifelse(ukb_data$PTVburden_app3 == 0, 0, 1))
tempdf = ukb_data %>% dplyr::select(eid, time, status, PTVburden_app3G, sex, s.color, all_of(sel_gpcas))
table(tempdf$status, tempdf$PTVburden_app3G, tempdf$sex)
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(tempdf)[4:ncol(tempdf)], collapse = "+")))
res.cox <- coxph(myformula, data = tempdf, id = tempdf$eid)
forest_model(model = res.cox) #1.1

#Filtered gene set - 71 genes
ensgsFilt = ensgs[colSums(mat_app2[ukb_data$status == 2,])>0] #genes with PTV occur in melanoma patients, 71 genes
ukb_data$PTVburden_app4 = apply(ukb_data[,ensgsFilt], 1, sum)
ukb_data$PTVburden_app4G = as.factor(ifelse(ukb_data$PTVburden_app4 == 0, 0, 1))
tempdf = ukb_data %>% dplyr::select(eid, time, status, PTVburden_app4G, sex, s.color, all_of(sel_gpcas))
table(tempdf$status, tempdf$PTVburden_app4G, tempdf$sex)
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(tempdf)[4:ncol(tempdf)], collapse = "+")))
res.cox <- coxph(myformula, data = tempdf, id = tempdf$eid)
forest_model(model = res.cox) #HR:1.28

```

```{r}
ukb_data_MEL = ukb_data %>% filter(status == 2)
View(cbind(geneID = genes_all$SYMBOL[fmatch(ensgsFilt, genes_all$ENSEMBL)], 
           female = colSums(ukb_data_MEL[ukb_data_MEL$sex == "female",ensgsFilt]), 
           male = colSums(ukb_data_MEL[ukb_data_MEL$sex == "male",ensgsFilt])))

```

#Finding of the best gene set

```{r}
load("Objects/go_desc_list")
ensgs_all_app = grep("antigen processing", names(go_desc_list), value = T)
go_desc_list = go_desc_list[ensgs_all_app]
hugos = unique(unlist(go_desc_list, use.names = F))
ensgs = genes_all$ENSEMBL[fmatch(hugos, genes_all$SYMBOL)]
ensgs = ensgs[!is.na(ensgs)] #108 genes

mat = pbsapply(ensgs, function(genes) {
  grepl(genes, ukb_data$PTVgenes)
})
colSums(mat)
ukb_data = cbind(ukb_data, mat)
rm(mat, go_desc_list, ensgs_all_app)

myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(ukb_data)[c(3,9:29,33:ncol(ukb_data))], collapse = "+")))
res.cox <- coxph(myformula, data = ukb_data, id = ukb_data$eid)
resSum = summary(res.cox)

#ENSGS with significant effect
ensgsSign = grep("ENSG", rownames(resSum$coefficients)[resSum$coefficients[,"Pr(>|z|)"] < 0.2], value = T)
ensgsSign = gsub("TRUE", "", ensgsSign)

ukb_data$PTVburden = apply(ukb_data[,ensgsSign], 1, sum)
ukb_data$PTVburdenG = as.factor(ifelse(ukb_data$PTVburden == 0, 0, 1))
tempdf = ukb_data %>% dplyr::select(eid, time, status, PTVburdenG, sex, s.color, all_of(sel_gpcas))
table(tempdf$status, tempdf$PTVburdenG, tempdf$sex)
myformula = as.formula(paste('Surv(time, status) ~ ', paste0(colnames(tempdf)[4:ncol(tempdf)], collapse = "+")))
coxph(myformula, data = tempdf, id = tempdf$eid)
forest_model(model = coxph(myformula, data = tempdf, id = tempdf$eid)) #HR:2.64



```


