---
title: "Collect frequent PTVs by tumors"
output: html_document
date: "2024-10-11"
---

#Setup

```{r}
setwd("/media/balazs/WorkL/balazs/Work/UKBiobank/")
Packages <- c("tidyverse", "magrittr", "org.Hs.eg.db", "clusterProfiler", "data.table", "future.apply", "Rfast", "fastmatch", "pbapply", "ggpubr")

library(survival, lib.loc = "/home/balazs/R/x86_64-pc-linux-gnu-library/4.3")

#, "h2o", , , "protr", , , "networkD3", "ggpubr", "RColorBrewer", "Biostrings", , "ggrepel", , "tidycmprsk", , "survminer", "ggsurvfit", "fastmatch", , "ComplexHeatmap", , "forestmodel", , )
lapply(Packages, require, character.only = TRUE)
rm(Packages)
options(dplyr.summarise.inform = FALSE)
```

#PTV gene frequencies

```{r}
ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

tsg = fread("Objects/tumor_supressor_genes_bioinfo_uth.txt")

ukb_data_unique = ukb_data %>% 
  dplyr::select(eid, Genes_PTV_MAF) %>% 
  unique()

genes_all = sort(unique(unlist(strsplit(ukb_data_unique$Genes_PTV_MAF, ","), use.names = F)))
genesFreq = bitr(geneID = genes_all, fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
genesFreq$tsg = genesFreq$SYMBOL %in% tsg$GeneSymbol
rm(genes_all)

#All UKB
temp = Table(unlist(lapply(ukb_data_unique$Genes_PTV_MAF, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_UKB = nrow(ukb_data_unique)
genesFreq$allUKB = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
rm(temp, ukb_data_unique)

#######
tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]
# v1 = ukb_data %>%
#   filter(cancer_type == "No_cancer") %>%
#   pull(eid)
# v2 = ukb_data %>%
#   filter(cancer_type %in% tumors) %>%
#   pull(eid)
# length(intersect(v1, v2)) #0

#Not tumorous
nottumorous = ukb_data %>% 
  filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)

temp = Table(unlist(lapply(nottumorous$Genes_PTV_MAF, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_notTumorous = nrow(nottumorous)
genesFreq$notTumorous = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
genesFreq$notTumorous[is.na(genesFreq$notTumorous)] = 0
rm(temp, nottumorous)

#Tumorous
tumorous = ukb_data %>% 
  filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)
temp = Table(unlist(lapply(tumorous$Genes_PTV_MAF, function(g) {
  unique(strsplit(g, ",")[[1]])
})))
genesFreq$n_tumorous = nrow(tumorous)
genesFreq$tumorous = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
rm(temp, tumorous)
genesFreq$tumorous[is.na(genesFreq$tumorous)] = 0


#In tumors separately
tumor_dfs = pblapply(tumors, function(t) {
  ukb_data %>% 
    filter(cancer_type == t) %>% 
    arrange(diag_date) %>% 
    distinct(eid, .keep_all = T)
})
names(tumor_dfs) = tumors

plan(multisession(workers = 8))
tumor_freqmat = future_lapply(tumors, function(t) {
  temp = Table(unlist(lapply(tumor_dfs[[t]]$Genes_PTV_MAF, function(g) {
    unique(strsplit(g, ",")[[1]])
  })))
  temp2 = unname(temp[fmatch(genesFreq$ENSEMBL, names(temp))])
  temp2[is.na(temp2)] = 0
  out = cbind(rep(nrow(tumor_dfs[[t]]), length(temp2)), temp2)
  colnames(out) = c(paste0("n_", t), t)
  out
})
tumor_freqmat2 = do.call(cbind, tumor_freqmat)
genesFreq = cbind(genesFreq, tumor_freqmat2)

saveRDS(genesFreq, file = "Objects/genesFreqMAF.rds")

```



med: 400, q3: 700 (Source: https://genomebiology.biomedcentral.com/articles/10.1186/s13059-023-02973-2)
MYO5A - 1,855 aas
MUC19 - 8,384 aas
RP1L1 - 2,400 aas
TTN - 34,350 aas
FLG - 4,061 aas
OBSCN - 7,968 aas
MUC16 - 14,507 aas
ITGB1BP1 - 200 aas !
ZNF90 - 601 aas



#I. Overrepresented tsgs in all tumors - Fisher tests

```{r}
genesFreq = readRDS("Objects/genesFreqMAF.rds")
genesFreq_TSG = genesFreq %>% filter(tsg==T)
plan(multisession(workers = 8))
ftpOR = t(future_apply(genesFreq_TSG, 1, function(x) {
  ft = fisher.test(rbind(c(as.numeric(x[9]), as.numeric(x[7])), c(as.numeric(x[8])-as.numeric(x[9]), as.numeric(x[6])-as.numeric(x[7]))))
  c(unname(ft$estimate), ft$p.value)
}))
ftpOR %<>% as.data.frame() %>% set_colnames(c("OR", "P")) %>% mutate(Padj = p.adjust(p = P, method = "BH"))
genesFreq_TSG = cbind(genesFreq_TSG[,1:9], ftpOR, genesFreq_TSG[,10:ncol(genesFreq_TSG)])

overrep_tsgs_ensg = genesFreq_TSG %>% filter(Padj<0.05) %>% pull(ENSEMBL)
overrep_tsgs_hugo = genesFreq_TSG %>% filter(Padj<0.05) %>% pull(SYMBOL)

save(overrep_tsgs_ensg, overrep_tsgs_hugo, file = "Objects/overrep_ft_tsgs_alltumors")

```

#II. Overrepresented TSGs by tumor types - Fisher tests

```{r}
genesFreq = readRDS("Objects/genesFreqMAF.rds")

ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]

n_nottumorous = ukb_data %>% 
  filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T) %>% 
  nrow() #354,783

tumor_patients_n = data.frame(cancer_type = tumors, 
                              n_patients = unname(sapply(tumors, function(t) {
                                ukb_data %>% 
                                  filter(cancer_type == t) %>% 
                                  arrange(diag_date) %>% 
                                  distinct(eid, .keep_all = T) %>% 
                                  nrow()
                              })))

genesFreq_TSG = genesFreq %>% filter(tsg==T)

plan(multisession(workers = 8))
overrep_genes_list = future_lapply(tumors, function(t) {
  ind = fmatch(t, colnames(genesFreq_TSG))
  n_tumorous = tumor_patients_n$n_patients[tumor_patients_n$cancer_type == t]
  ftpOR = t(apply(genesFreq_TSG, 1, function(x) {
    ft = fisher.test(rbind(c(as.numeric(x[ind]), as.numeric(x[5])), c(n_tumorous-as.numeric(x[ind]), n_nottumorous-as.numeric(x[5]))))
    c(unname(ft$estimate), ft$p.value)
  }))
  ftpOR %<>% as.data.frame() %>% set_colnames(c("OR", "P")) %>% mutate(Padj = p.adjust(p = P, method = "BH"))
  cbind(genesFreq_TSG[,c("ENSEMBL", "SYMBOL", "allUKB", "notTumorous", "tumorous")], ftpOR)
})
names(overrep_genes_list) = tumors
overrep_tsgs_ensg = lapply(tumors, function(t) overrep_genes_list[[t]] %>% filter(Padj<0.05) %>% pull(ENSEMBL))
names(overrep_tsgs_ensg) = tumors
overrep_tsgs_hugo = lapply(tumors, function(t) overrep_genes_list[[t]] %>% filter(Padj<0.05) %>% pull(SYMBOL))
names(overrep_tsgs_hugo) = tumors
lengths(overrep_tsgs_ensg)

save(overrep_tsgs_ensg, overrep_tsgs_hugo, file = "Objects/overrep_ft_tsgs_tumors")

```

#III. Overrepresented TSGs by tumor types and gender - Fisher tests

```{r}
ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]

tumor_freq_data = as.data.frame(table(ukb_data$cancer_type, ukb_data$sex)) %>%
  set_colnames(c("cancer_type", "sex", "n")) %>% 
  mutate_if(is.factor, as.character) %>% 
  filter(cancer_type %in% tumors) %>% 
  filter(n >= 20)

tsg = fread("Objects/tumor_supressor_genes_bioinfo_uth.txt")

```



#Compare the frequency of PTVs in TSG and other genes

```{r}
genesFreq = readRDS("Objects/genesFreqMAF.rds")

genesFreq %<>% mutate(ratio_T_NT = tumorous/notTumorous)
quantile(genesFreq$ratio_T_NT, na.rm = T)

genesFreq %>% 
  ggplot(aes(x = tsg, y = ratio_T_NT)) + 
  geom_boxplot() + 
  stat_compare_means() + 
  scale_y_continuous(trans = "log2")

genesFreq %>% 
  filter(tumorous > 100) %>% 
  ggplot(aes(x = tsg, y = ratio_T_NT)) + 
  geom_boxplot() + 
  stat_compare_means() + 
  scale_y_continuous(trans = "log2")
```

#Overrepresented genes in different tumors

```{r}
genesFreq = readRDS("Objects/genesFreqMAF.rds")

ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]

n_nottumorous = ukb_data %>% 
  filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T) %>% 
  nrow() #354,783

tumor_patients_n = data.frame(cancer_type = tumors, 
                              n_patients = unname(sapply(tumors, function(t) {
                                ukb_data %>% 
                                  filter(cancer_type == t) %>% 
                                  arrange(diag_date) %>% 
                                  distinct(eid, .keep_all = T) %>% 
                                  nrow()
                              })))

temp = sapply(tumors, function(t) {
  genesFreq[,t]/tumor_patients_n$n_patients[tumor_patients_n$cancer_type == t]
})
colnames(temp) = paste0(tumors, "_R")
notTumorous_R = genesFreq$notTumorous/n_nottumorous
genesFreq = cbind(genesFreq, notTumorous_R, temp)
rm(temp, notTumorous_R)

temp = sapply(tumors, function(t) {
  genesFreq[,paste0(t, "_R")] / genesFreq[,"notTumorous_R"]
})
colnames(temp) = paste0(tumors, "RPerNTR")
genesFreq = cbind(genesFreq, temp)
rm(temp)

overrep_genes_tumors = lapply(tumors, function(t) {
  tempdf = genesFreq[,c("ENSEMBL", "SYMBOL", t, paste0(t, "RPerNTR"))]
  colnames(tempdf)[3:4] = c("Tumor", "TumorRPerNTR")
  n1perc = round(tumor_patients_n$n_patients[tumor_patients_n$cancer_type == t]/1000)
  tempdf %<>% filter(Tumor > n1perc)
  ensgs = tempdf %>% arrange(desc(TumorRPerNTR)) %>% pull(ENSEMBL)
  hugos = tempdf %>% arrange(desc(TumorRPerNTR)) %>% pull(SYMBOL)
  ensgs = ensgs[1:30]
  hugos = hugos[1:30]
  list(ensgs[!is.na(ensgs)], hugos[!is.na(hugos)])
  #rm(tempdf, ensgs, hugos, n1perc, t)
})
names(overrep_genes_tumors) = tumors

saveRDS(overrep_genes_tumors, file = "Objects/overrep_genes_tumors.rds")

#TSGs
genesFreq_TSG = genesFreq %>% filter(tsg == T)

overrep_tsgs_tumors = lapply(tumors, function(t) {
  tempdf = genesFreq_TSG[,c("ENSEMBL", "SYMBOL", t, paste0(t, "RPerNTR"))]
  colnames(tempdf)[3:4] = c("Tumor", "TumorRPerNTR")
  n1perc = round(tumor_patients_n$n_patients[tumor_patients_n$cancer_type == t]/1000)
  tempdf %<>% filter(Tumor > n1perc)
  ensgs = tempdf %>% arrange(desc(TumorRPerNTR)) %>% pull(ENSEMBL)
  hugos = tempdf %>% arrange(desc(TumorRPerNTR)) %>% pull(SYMBOL)
  ensgs = ensgs[1:30]
  hugos = hugos[1:30]
  list(ensgs[!is.na(ensgs)], hugos[!is.na(hugos)])
  #rm(tempdf, ensgs, hugos, n1perc, t)
})
names(overrep_tsgs_tumors) = tumors

saveRDS(overrep_tsgs_tumors, file = "Objects/overrep_tsgs_tumors.rds")

```

#Overrepresented genes in all tumors - Fisher tests

```{r}
genesFreq = readRDS("Objects/genesFreqMAF.rds")

ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))

tumors = sort(unique(ukb_data$cancer_type))
tumors = tumors[!(tumors %in% c("Benign", "Hydatidiform_mole", "No_cancer", "other", "undef"))]

n_nottumorous = ukb_data %>% 
  filter(cancer_type == "No_cancer") %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T) %>% 
  nrow() #354,783

tumorous = ukb_data %>% 
  filter(cancer_type %in% tumors) %>% 
  arrange(diag_date) %>% 
  distinct(eid, .keep_all = T)

```


