---
title: "B_Incidence_age_sex_PTV"
author: "Balazs Koncz"
date: "2024-07-30"
output: html_document
---

#Setup

```{r}
library(dplyr)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(fastmatch)
options(dplyr.summarise.inform = FALSE)
```

#PTV_universal_1

```{r}
load("objects/ukb_data_cancer_final")
ukb_data_cancer$cancer_type[is.na(ukb_data_cancer$cancer_type)] = "No_cancer"
tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

age_inc_tbl = lapply(X = tumors, FUN = function(x){
  tempdf_x = ukb_data_cancer %>% filter(cancer_type == x) %>% arrange(diag_date)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
  tempdf_x %<>% mutate(inc_time = round(as.numeric((diag_date - birthdate))/365.25),
                       final_age = case_when(
                         death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                         death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)))
  
  tempdf_no_x = ukb_data_cancer %>% filter(is.na(ICD10) | substr(ICD10,1,1) != "C")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
  tempdf_no_x %<>% mutate(inc_time = NA, 
                          final_age = case_when(
                            death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                            death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)))
  
  minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
  maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))
  
  res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data_cancer$sex), PTV_universal_1 = levels(tempdf_x$PTV_universal_1), stringsAsFactors = F) 
  
  res_case = tempdf_x %>% group_by(inc_time, sex, PTV_universal_1) %>% summarise(case = n())

  res %<>% left_join(res_case)
  res$case[is.na(res$case)] = 0
  
  res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_universal_1) %>% summarise(no_case = n())


  res$no_case = 0
  for(i in 1:nrow(res_nocase)) {
    res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_universal_1 == res_nocase$PTV_universal_1[i]] = res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_universal_1 == res_nocase$PTV_universal_1[i]] + res_nocase$no_case[i]
  }
  
  res$PTV_universal_1 = factor(res$PTV_universal_1, levels = levels(tempdf_x$PTV_universal_1))
  res %<>% mutate(inc = case/(case+no_case)) %>% mutate(tumor = x, .before = "inc_time")
  res
})
age_inc_tbl = bind_rows(age_inc_tbl)

age_inc_plots = lapply(X = tumors, FUN = function(x){
  res = age_inc_tbl %>% filter(tumor == x)
  ggplot(res, mapping = aes(x = inc_time, y = inc, color = PTV_universal_1, shape = as.factor(sex))) + 
    geom_point() + 
    geom_smooth() +
    facet_wrap(~sex) + 
    labs(title = x)
})

ggarrange(plotlist = age_inc_plots, ncol = 5, nrow = 5)
ggsave(filename = "plots/age_inc_plot_sex_PTV_universal_1.jpg", width = 100, height = 60, units = "cm", dpi = 300)
```

#Age intervals

```{r}
age_inc_tbl2 = lapply(X = tumors, FUN = function(x){
  tempdf = age_inc_tbl %>% filter(tumor == x)
  res = expand.grid(inc_time_int = c("10_15", "16_20", "21_25", "26_30", "31_35", "36_40", "41_45", "46_50", "51_55", "56_60", "61_65", "66_70", "71_75", "76_80", "81_85", "86_90"),
                     sex = unique(ukb_data_cancer$sex),
                     PTV_universal_1 = levels(tempdf$PTV_universal_1), stringsAsFactors = F)
  
  res$case = apply(res, 1, function(y) {
    tdf = tempdf %>% filter(inc_time >= as.numeric(substr(y[1],1,2)), inc_time <= as.numeric(substr(y[1],4,5)), sex == as.numeric(y[2]), PTV_universal_1 == y[3])
    sum(tdf$case)
  })
  res$no_case = apply(res, 1, function(y) {
    tdf = tempdf %>% filter(inc_time >= as.numeric(substr(y[1],1,2)), inc_time <= as.numeric(substr(y[1],4,5)), sex == as.numeric(y[2]), PTV_universal_1 == y[3])
    if(nrow(tdf) == 0) {
      out = 0
    } else {
      out = tdf$no_case[nrow(tdf)]
    }
    out
  })
  res$PTV_universal_1 = factor(res$PTV_universal_1, levels = levels(tempdf$PTV_universal_1))
  res %<>% mutate(inc = case/(case+no_case)) %>% mutate(tumor = x, .before = "inc_time_int")
  res$inc_time_int = factor(res$inc_time_int, levels = c("10_15", "16_20", "21_25", "26_30", "31_35", "36_40", "41_45", "46_50", "51_55", "56_60", "61_65", "66_70", "71_75", "76_80", "81_85", "86_90"))
  res
})

age_inc_tbl2 = bind_rows(age_inc_tbl2)

age_inc_plots2 = lapply(X = tumors, FUN = function(x){
  res = age_inc_tbl2 %>% filter(tumor == x)
  ggplot(res, mapping = aes(x = inc_time_int, y = inc, color = PTV_universal_1, shape = as.factor(sex))) + 
    geom_point() + 
    geom_smooth() +
    facet_wrap(~sex) + 
    labs(title = x)
})

ggarrange(plotlist = age_inc_plots2, ncol = 5, nrow = 5)
ggsave(filename = "plots/age_int_inc_plot_sex_PTV_universal_1.jpg", width = 100, height = 60, units = "cm", dpi = 300)
```

#Tumor incidence in patients with the highest PTV burden

```{r}
load("objects/ukb_data_cancer_final")
ukb_data_cancer$cancer_type[is.na(ukb_data_cancer$cancer_type)] = "No_cancer"
tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]
qs = quantile(ukb_data_cancer$PTV_burden, seq(0.1,1,0.05))
cutoffs = qs[c("75%","80%","85%","90%","95%")]

age_inc_tbl_co = lapply(cutoffs, function(co) {
  age_inc_tbl = lapply(X = tumors, FUN = function(x){
    tempdf_x = ukb_data_cancer %>% filter(cancer_type == x) %>% arrange(diag_date)
    tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
    tempdf_x %<>% mutate(inc_time = round(as.numeric((diag_date - birthdate))/365.25),
                         final_age = case_when(
                           death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                           death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                         PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))
    
    tempdf_no_x = ukb_data_cancer %>% filter(is.na(ICD10) | substr(ICD10,1,1) != "C")
    tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
    tempdf_no_x %<>% mutate(inc_time = NA, 
                            final_age = case_when(
                              death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                              death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                            PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))
    
    minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    
    res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data_cancer$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 
    
    res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n())
    res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
    res$case[is.na(res$case)] = 0
    
    res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())
    
    res$no_case = 0
    for(i in 1:nrow(res_nocase)) {
      res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] = res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] + res_nocase$no_case[i]
    }
    
    res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
    res %<>% 
      mutate(inc = case/(case+no_case)) %>% 
      mutate(tumor = x, .before = "inc_time") %>% 
      mutate(PTVcutoff = co, .before = "inc_time")
    res$inc[is.nan(res$inc)] = 0
    res
  })
  age_inc_tbl = bind_rows(age_inc_tbl)
  age_inc_tbl
})


lapply(1:length(age_inc_tbl_co), function(i) {
  age_inc_plots = lapply(X = tumors, FUN = function(x) {
    tempdf = age_inc_tbl_co[[i]] %>% filter(tumor == x)
    ggplot(tempdf, mapping = aes(x = inc_time, y = inc, color = PTV_burden_group, shape = as.factor(sex))) + 
      geom_point() + 
      geom_smooth() +
      facet_wrap(~sex) + 
      labs(title = x)
  })
  ggarrange(plotlist = age_inc_plots, ncol = 5, nrow = 5)
  ggsave(filename = paste0("plots/age_inc_plot_sex_PTVco", cutoffs[i] ,".jpg"), width = 100, height = 60, units = "cm", dpi = 300)
})


```

#modified PTV burden - immune, tsg, pLI
##Tumor incidence in patients with the highest immune PTV burden

```{r}
load("objects/ukb_data_cancer_final")
load("objects/ptvb_immune")

ukb_data_cancer$cancer_type[is.na(ukb_data_cancer$cancer_type)] = "No_cancer"
tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

ukb_data_cancer$PTV_burden = ptvb$PTVim[fmatch(as.character(ukb_data_cancer$eid), ptvb$Patient.ID)]

qs = quantile(ukb_data_cancer$PTV_burden, seq(0.1,1,0.05))
cutoffs = unique(qs[c("75%","80%","85%","90%","95%")])

age_inc_tbl_co = lapply(cutoffs, function(co) {
  age_inc_tbl = lapply(X = tumors, FUN = function(x){
    tempdf_x = ukb_data_cancer %>% filter(cancer_type == x) %>% arrange(diag_date)
    tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
    tempdf_x %<>% mutate(inc_time = round(as.numeric((diag_date - birthdate))/365.25),
                         final_age = case_when(
                           death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                           death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                         PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))
    
    tempdf_no_x = ukb_data_cancer %>% filter(is.na(ICD10) | !substr(ICD10,1,1) %in% c("C", "D"))
    tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
    tempdf_no_x %<>% mutate(inc_time = NA, 
                            final_age = case_when(
                              death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                              death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                            PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))
    
    minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    
    res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data_cancer$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 
    
    res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n())
    res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
    res$case[is.na(res$case)] = 0
    
    res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())
    
    res$no_case = 0
    for(i in 1:nrow(res_nocase)) {
      res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] = res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] + res_nocase$no_case[i]
    }
    
    res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
    res %<>% 
      mutate(inc = case/(case+no_case)) %>% 
      mutate(tumor = x, .before = "inc_time") %>% 
      mutate(PTVcutoff = co, .before = "inc_time")
    res$inc[is.nan(res$inc)] = 0
    res
  })
  age_inc_tbl = bind_rows(age_inc_tbl)
  age_inc_tbl
})


lapply(1:length(age_inc_tbl_co), function(i) {
  age_inc_plots = lapply(X = tumors, FUN = function(x) {
    tempdf = age_inc_tbl_co[[i]] %>% filter(tumor == x)
    ggplot(tempdf, mapping = aes(x = inc_time, y = inc, color = PTV_burden_group, shape = as.factor(sex))) + 
      geom_point() + 
      geom_smooth() +
      facet_wrap(~sex) + 
      labs(title = x)
  })
  ggarrange(plotlist = age_inc_plots, ncol = 5, nrow = 5)
  ggsave(filename = paste0("plots/age_inc_plot_sex_imPTVco", cutoffs[i] ,".jpg"), width = 100, height = 60, units = "cm", dpi = 300)
})


```

##Tumor incidence

```{r}
load("objects/ukb_data_cancer_final")
load("objects/ptvb_immune")

ukb_data_cancer$cancer_type[is.na(ukb_data_cancer$cancer_type)] = "No_cancer"
tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

ukb_data_cancer$PTV_burden = ptvb$PTVim[fmatch(as.character(ukb_data_cancer$eid), ptvb$Patient.ID)]
hist(ukb_data_cancer$PTV_burden)
qs = quantile(ukb_data_cancer$PTV_burden, seq(0.1,1,0.05))
cutoffs = unique(qs[c("75%","80%","85%","90%","95%")])
co = 2


tempdf_x = ukb_data_cancer %>% filter(substr(ICD10,1,1) == "C") %>% arrange(diag_date)
tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
tempdf_x %<>% mutate(inc_time = round(as.numeric((diag_date - birthdate))/365.25),
                     final_age = case_when(
                       death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                       death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                     PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))

tempdf_no_x = ukb_data_cancer %>% filter(is.na(ICD10) | !substr(ICD10,1,1) %in% c("C", "D"))
tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
tempdf_no_x %<>% mutate(inc_time = NA, 
                        final_age = case_when(
                          death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                          death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                        PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))

minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))

res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data_cancer$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 

res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n())
res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
res$case[is.na(res$case)] = 0

res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())

res$no_case = 0
for(i in 1:nrow(res_nocase)) {
  res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] = res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] + res_nocase$no_case[i]
}

res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
res %<>% 
  mutate(inc = case/(case+no_case)) %>% 
  mutate(tumor = "any", .before = "inc_time") %>% 
  mutate(PTVcutoff = co, .before = "inc_time")
res$inc[is.nan(res$inc)] = 0

ggplot(res, mapping = aes(x = inc_time, y = inc, color = PTV_burden_group, shape = as.factor(sex))) + 
  geom_point() + 
  geom_smooth(method = "loess") +
  facet_wrap(~sex) + 
  labs(title = "Tumor incidence")

ggsave(filename = paste0("plots/age_inc_plot_sex_imPTVco", co ,"any_tumor.jpg"), width = 30, height = 25, units = "cm", dpi = 300)
```



#PTV - Tumor Suppressor genes
https://bioinfo.uth.edu/TSGene/download.cgi?csrt=15728982272603520603
##Tumor incidence in patients with the highest tsg PTV burden

```{r}
load("objects/ukb_data_cancer_final")
load("objects/ptvb_tsg")

ukb_data_cancer$cancer_type[is.na(ukb_data_cancer$cancer_type)] = "No_cancer"
tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

ukb_data_cancer$PTV_burden = ptvb$PTVtsg[fmatch(as.character(ukb_data_cancer$eid), ptvb$Patient.ID)]
hist(ukb_data_cancer$PTV_burden)
qs = quantile(ukb_data_cancer$PTV_burden, seq(0.1,1,0.05))
cutoffs = unique(qs[c("75%","80%","85%","90%","95%")])

age_inc_tbl_co = lapply(cutoffs, function(co) {
  age_inc_tbl = lapply(X = tumors, FUN = function(x){
    tempdf_x = ukb_data_cancer %>% filter(cancer_type == x) %>% arrange(diag_date)
    tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
    tempdf_x %<>% mutate(inc_time = round(as.numeric((diag_date - birthdate))/365.25),
                         final_age = case_when(
                           death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                           death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                         PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))
    
    tempdf_no_x = ukb_data_cancer %>% filter(is.na(ICD10) | !substr(ICD10,1,1) %in% c("C", "D"))
    tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
    tempdf_no_x %<>% mutate(inc_time = NA, 
                            final_age = case_when(
                              death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                              death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                            PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))
    
    minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    
    res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data_cancer$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 
    
    res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n())
    res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
    res$case[is.na(res$case)] = 0
    
    res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())
    
    res$no_case = 0
    for(i in 1:nrow(res_nocase)) {
      res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] = res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] + res_nocase$no_case[i]
    }
    
    res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
    res %<>% 
      mutate(inc = case/(case+no_case)) %>% 
      mutate(tumor = x, .before = "inc_time") %>% 
      mutate(PTVcutoff = co, .before = "inc_time")
    res$inc[is.nan(res$inc)] = 0
    res
  })
  age_inc_tbl = bind_rows(age_inc_tbl)
  age_inc_tbl
})



lapply(1:length(age_inc_tbl_co), function(i) {
  age_inc_plots = lapply(X = tumors, FUN = function(x) {
    tempdf = age_inc_tbl_co[[i]] %>% filter(tumor == x)
    ggplot(tempdf, mapping = aes(x = inc_time, y = inc, color = PTV_burden_group, shape = as.factor(sex))) + 
      geom_point() + 
      geom_smooth() +
      facet_wrap(~sex) + 
      labs(title = x)
  })
  ggarrange(plotlist = age_inc_plots, ncol = 5, nrow = 5)
  ggsave(filename = paste0("plots/age_inc_plot_sex_tsgPTVco", cutoffs[i] ,".jpg"), width = 100, height = 60, units = "cm", dpi = 300)
})



```

##Tumor incidence

```{r}
load("objects/ukb_data_cancer_final")
load("objects/ptvb_tsg")

ukb_data_cancer$cancer_type[is.na(ukb_data_cancer$cancer_type)] = "No_cancer"
tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

ukb_data_cancer$PTV_burden = ptvb$PTVtsg[fmatch(as.character(ukb_data_cancer$eid), ptvb$Patient.ID)]
hist(ukb_data_cancer$PTV_burden)
qs = quantile(ukb_data_cancer$PTV_burden, seq(0.1,1,0.05))
cutoffs = unique(qs[c("75%","80%","85%","90%","95%")])
co = 3


tempdf_x = ukb_data_cancer %>% filter(substr(ICD10,1,1) == "C") %>% arrange(diag_date)
tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
tempdf_x %<>% mutate(inc_time = round(as.numeric((diag_date - birthdate))/365.25),
                     final_age = case_when(
                       death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                       death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                     PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))

tempdf_no_x = ukb_data_cancer %>% filter(is.na(ICD10) | !substr(ICD10,1,1) %in% c("C", "D"))
tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
tempdf_no_x %<>% mutate(inc_time = NA, 
                        final_age = case_when(
                          death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                          death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                        PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(qs)+1), right = F))

minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))

res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data_cancer$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 

res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n())
res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
res$case[is.na(res$case)] = 0

res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())

res$no_case = 0
for(i in 1:nrow(res_nocase)) {
  res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] = res$no_case[res$inc_time <= res_nocase$final_age[i] & res$sex == res_nocase$sex[i] & res$PTV_burden_group == res_nocase$PTV_burden_group[i]] + res_nocase$no_case[i]
}

res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
res %<>% 
  mutate(inc = case/(case+no_case)) %>% 
  mutate(tumor = "any", .before = "inc_time") %>% 
  mutate(PTVcutoff = co, .before = "inc_time")
res$inc[is.nan(res$inc)] = 0

ggplot(res, mapping = aes(x = inc_time, y = inc, color = PTV_burden_group, shape = as.factor(sex))) + 
  geom_point() + 
  geom_smooth(method = "loess") +
  facet_wrap(~sex) + 
  labs(title = "Tumor incidence")

ggsave(filename = paste0("plots/age_inc_plot_sex_tsgPTVco", co ,"any_tumor.jpg"), width = 30, height = 25, units = "cm", dpi = 300)
```

#Tumor incidence in patients in a given age with the highest immune PTV burden - more metrics (immune, tsg, pLI), more cutoffs 
##Per tumor types

MISTAKE - higher cutoff, lower curve

```{r}
load("objects/ukb_data_cancer_final")
load("objects/ptvb_im_tsg_pli")
colnames(ptvb)[1] = "eid"
ptvb$eid = as.numeric(ptvb$eid)

tumors = sort(unique(ukb_data_cancer$cancer_type))
tumors = tumors[tumors != "No_cancer"]

colnames(ptvb)[3] = "PTVb"
ptv_burden_metrics = colnames(ptvb)[c(3,7:17)]

for(i in 1:length(ptv_burden_metrics)) {
  print(i)
  ukb_data = ukb_data_cancer[,c("eid", "sex", "birthdate", "ICD10", "cancer_type", "diag_date", "death", "date_of_death")]
  ptvb_temp = ptvb %>% select("eid", ptv_burden_metrics[i])
  colnames(ptvb_temp) = c("eid", "PTV_burden")
  ukb_data$PTV_burden = ptvb_temp$PTV_burden[fmatch(ukb_data$eid, ptvb_temp$eid)]
  cutoffs = unique(unname(quantile(ukb_data$PTV_burden, seq(0.75,0.95,0.05))))
  if(all(cutoffs == 0)) {cutoffs = 1}
  lapply(cutoffs, function(co) {
    age_inc_tbl = lapply(X = tumors, FUN = function(x) {
      tempdf_x = ukb_data %>% filter(cancer_type == x) %>% arrange(diag_date)
      tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
      tempdf_x %<>% mutate(inc_time = floor(as.numeric((diag_date - birthdate))/365.25),
                           final_age = case_when(
                             death == T ~ floor(as.numeric(date_of_death - birthdate)/365.25), #FLOOR!!!!!!!!!!!!!!!
                             death == F ~ floor(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                           PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(ukb_data$PTV_burden)+1), right = F))
      
      tempdf_no_x = ukb_data %>% filter(!eid %fin% tempdf_x$eid)
      tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
      tempdf_no_x %<>% mutate(inc_time = NA, 
                              final_age = case_when(
                                death == T ~ floor(as.numeric(date_of_death - birthdate)/365.25),
                                death == F ~ floor(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                              PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(ukb_data$PTV_burden)+1), right = F))
      
      minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
      maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))
      res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F)
      #Patients with a given tumor
      res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n()) #age at tumor diagnosis
      res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
      res$case[is.na(res$case)] = 0
      
      res_case2 = tempdf_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(case2 = n()) #final age of patients with a given tumor
      res$case2 = 0
      for(k in 1:nrow(res_case2)) {
        res$case2[res$inc_time <= res_case2$final_age[k] & res$sex == res_case2$sex[k] & res$PTV_burden_group == res_case2$PTV_burden_group[k]] = res$case2[res$inc_time <= res_case2$final_age[k] & res$sex == res_case2$sex[k] & res$PTV_burden_group == res_case2$PTV_burden_group[k]] + res_case2$case2[k]
      }
      #Patients without the given tumor
      res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())
      res$no_case = 0
      for(j in 1:nrow(res_nocase)) {
        res$no_case[res$inc_time <= res_nocase$final_age[j] & res$sex == res_nocase$sex[j] & res$PTV_burden_group == res_nocase$PTV_burden_group[j]] = res$no_case[res$inc_time <= res_nocase$final_age[j] & res$sex == res_nocase$sex[j] & res$PTV_burden_group == res_nocase$PTV_burden_group[j]] + res_nocase$no_case[j]
      }
      res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
      res %<>% 
        mutate(inc = 100000*case/(case2+no_case)) %>% 
        mutate(tumor = x, .before = "inc_time") %>% 
        mutate(PTVcutoff = co, .before = "inc_time")
      res$inc[is.nan(res$inc)] = 0
      res
    })
    age_inc_tbl = bind_rows(age_inc_tbl)
    age_inc_plots = lapply(X = tumors, FUN = function(z) {
      tempdf = age_inc_tbl %>% filter(tumor == z)
      w0 = wilcox.test(inc ~ PTV_burden_group, tempdf[tempdf$sex == 0,])$p.value
      w1 = wilcox.test(inc ~ PTV_burden_group, tempdf[tempdf$sex == 1,])$p.value
      ggarrange(plotlist = list(
        ggplot(tempdf[tempdf$sex == 0,], mapping = aes(x = inc_time, y = inc, color = PTV_burden_group)) + 
          geom_point() + 
          geom_smooth() +
          labs(title = paste0(z, " Women P = ", round(w0,4))),
        ggplot(tempdf[tempdf$sex == 1,], mapping = aes(x = inc_time, y = inc, color = PTV_burden_group)) + 
          geom_point() + 
          geom_smooth() +
          labs(title = paste0(z, " Men P = ", round(w1,4)))
      ), ncol = 2, common.legend = T)
    })
    fig1 = ggarrange(plotlist = age_inc_plots, ncol = 5, nrow = 5, common.legend = T)
    ggsave(filename = paste0("plots/tumor_incidence/tumor_incidence_", ptv_burden_metrics[i], "_cutoff_", co ,".jpg"), plot = fig1, width = 100, height = 60, units = "cm", dpi = 300)
  })
}

```

##All tumor incidence

```{r}
load("objects/ukb_data_cancer_final")
load("objects/ptvb_im_tsg_pli")
colnames(ptvb)[1] = "eid"
ptvb$eid = as.numeric(ptvb$eid)

tumors = sort(unique(ukb_data_cancer$cancer_type))
tumors = tumors[tumors != "No_cancer"]

colnames(ptvb)[3] = "PTVb"
ptv_burden_metrics = colnames(ptvb)[c(3,7:17)]

for(i in 1:length(ptv_burden_metrics)) {
  print(i)
  ukb_data = ukb_data_cancer[,c("eid", "sex", "birthdate", "ICD10", "cancer_type", "diag_date", "death", "date_of_death")]
  ptvb_temp = ptvb %>% select("eid", ptv_burden_metrics[i])
  colnames(ptvb_temp) = c("eid", "PTV_burden")
  ukb_data$PTV_burden = ptvb_temp$PTV_burden[fmatch(ukb_data$eid, ptvb_temp$eid)]
  cutoffs = unique(unname(quantile(ukb_data$PTV_burden, seq(0.75,0.95,0.05))))
  if(all(cutoffs == 0)) {cutoffs = 1}
  lapply(cutoffs, function(co) {
    tempdf_x = ukb_data %>% filter(!is.na(ICD10)) %>% arrange(diag_date)
    tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
    tempdf_x %<>% mutate(inc_time = floor(as.numeric((diag_date - birthdate))/365.25),
                         final_age = case_when(
                           death == T ~ floor(as.numeric(date_of_death - birthdate)/365.25),
                           death == F ~ floor(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                         PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(ukb_data$PTV_burden)+1), right = F))
    tempdf_no_x = ukb_data %>% filter(is.na(ICD10))
    tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
    tempdf_no_x %<>% mutate(inc_time = NA, 
                            final_age = case_when(
                              death == T ~ floor(as.numeric(date_of_death - birthdate)/365.25),
                              death == F ~ floor(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                            PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(ukb_data$PTV_burden)+1), right = F))
    minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))
    res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 
    #Patients with tumor
    res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n())
    res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
    res$case[is.na(res$case)] = 0
    
    res_case2 = tempdf_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(case2 = n()) #final age of patients with a given tumor
    res$case2 = 0
    for(k in 1:nrow(res_case2)) {
      res$case2[res$inc_time <= res_case2$final_age[k] & res$sex == res_case2$sex[k] & res$PTV_burden_group == res_case2$PTV_burden_group[k]] = res$case2[res$inc_time <= res_case2$final_age[k] & res$sex == res_case2$sex[k] & res$PTV_burden_group == res_case2$PTV_burden_group[k]] + res_case2$case2[k]
    }
    #Patients without tumor
    res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())
    res$no_case = 0
    for(j in 1:nrow(res_nocase)) {
      res$no_case[res$inc_time <= res_nocase$final_age[j] & res$sex == res_nocase$sex[j] & res$PTV_burden_group == res_nocase$PTV_burden_group[j]] = res$no_case[res$inc_time <= res_nocase$final_age[j] & res$sex == res_nocase$sex[j] & res$PTV_burden_group == res_nocase$PTV_burden_group[j]] + res_nocase$no_case[j]
    }
    res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
    res %<>% 
      mutate(inc = 100000*case/(case2+no_case)) %>% 
      mutate(PTVcutoff = co, .before = "inc_time")
    res$inc[is.nan(res$inc)] = 0
    #Statistics
    w0 = wilcox.test(inc ~ PTV_burden_group, res[res$sex == 0,])$p.value
    w1 = wilcox.test(inc ~ PTV_burden_group, res[res$sex == 1,])$p.value
    #Plot
    fig1 = ggarrange(plotlist = list(
      ggplot(res[res$sex == 0,], mapping = aes(x = inc_time, y = inc, color = PTV_burden_group)) + 
        geom_point() + 
        geom_smooth() +
        labs(title = paste0("Tumor incidence (women) P = ", round(w0,4))),
      ggplot(res[res$sex == 1,], mapping = aes(x = inc_time, y = inc, color = PTV_burden_group)) + 
        geom_point() + 
        geom_smooth() +
        labs(title = paste0("Tumor incidence (men) P = ", round(w1,4)))
    ), ncol = 2, common.legend = T)
    ggsave(filename = paste0("plots/tumor_incidence/alltumor_incidence_", ptv_burden_metrics[i], "_cutoff_", co ,".jpg"), plot = fig1, width = 30, height = 25, units = "cm", dpi = 300)
  })
}
```

##Age intervals (per tumor)

```{r}
load("objects/ukb_data_cancer_final")
load("objects/ptvb_im_tsg_pli")
colnames(ptvb)[1] = "eid"
ptvb$eid = as.numeric(ptvb$eid)

tumors = sort(unique(ukb_data_cancer$cancer_type))
tumors = tumors[tumors != "No_cancer"]

colnames(ptvb)[3] = "PTVb"
ptv_burden_metrics = colnames(ptvb)[c(3,7:17)]

for(i in 1:length(ptv_burden_metrics)) {
  print(i)
  ukb_data = ukb_data_cancer[,c("eid", "sex", "birthdate", "ICD10", "cancer_type", "diag_date", "death", "date_of_death")]
  ptvb_temp = ptvb %>% select("eid", ptv_burden_metrics[i])
  colnames(ptvb_temp) = c("eid", "PTV_burden")
  ukb_data$PTV_burden = ptvb_temp$PTV_burden[fmatch(ukb_data$eid, ptvb_temp$eid)]
  cutoffs = unique(unname(quantile(ukb_data$PTV_burden, seq(0.75,0.95,0.05))))
  if(all(cutoffs == 0)) {cutoffs = 1}
  
  lapply(cutoffs, function(co) {
    age_inc_plots = lapply(X = tumors, FUN = function(x) {
      tempdf_x = ukb_data %>% filter(cancer_type == x) %>% arrange(diag_date)
      tempdf_x = tempdf_x[!duplicated(tempdf_x$eid),]
      tempdf_x %<>% mutate(inc_time = round(as.numeric((diag_date - birthdate))/365.25),
                           final_age = case_when(
                             death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                             death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                           PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(ukb_data$PTV_burden)+1), right = F))
      tempdf_no_x = ukb_data %>% filter(is.na(ICD10))
      tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid),]
      tempdf_no_x %<>% mutate(inc_time = NA, 
                              final_age = case_when(
                                death == T ~ round(as.numeric(date_of_death - birthdate)/365.25),
                                death == F ~ round(as.numeric(max(date_of_death, na.rm = T) - birthdate)/365.25)),
                              PTV_burden_group = cut(PTV_burden, breaks = c(0,co,max(ukb_data$PTV_burden)+1), right = F))
      
      minage = min(c(tempdf_x$inc_time, tempdf_no_x$final_age))
      maxage = max(c(tempdf_x$inc_time, tempdf_no_x$final_age))
      
      res = expand.grid(inc_time = seq(minage, maxage, 1), sex = unique(ukb_data$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 
      #Patients with tumor
      res_case = tempdf_x %>% group_by(inc_time, sex, PTV_burden_group) %>% summarise(case = n())
      res %<>% left_join(res_case, by = c("inc_time", "sex", "PTV_burden_group"))
      res$case[is.na(res$case)] = 0
      
      res_case2 = tempdf_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(case2 = n()) #final age of patients with a given tumor
      res$case2 = 0
      for(k in 1:nrow(res_case2)) {
        res$case2[res$inc_time <= res_case2$final_age[k] & res$sex == res_case2$sex[k] & res$PTV_burden_group == res_case2$PTV_burden_group[k]] = res$case2[res$inc_time <= res_case2$final_age[k] & res$sex == res_case2$sex[k] & res$PTV_burden_group == res_case2$PTV_burden_group[k]] + res_case2$case2[k]
      }
      #Patients without tumor
      res_nocase = tempdf_no_x %>% group_by(final_age, sex, PTV_burden_group) %>% summarise(no_case = n())
      res$no_case = 0
      for(j in 1:nrow(res_nocase)) {
        res$no_case[res$inc_time <= res_nocase$final_age[j] & res$sex == res_nocase$sex[j] & res$PTV_burden_group == res_nocase$PTV_burden_group[j]] = res$no_case[res$inc_time <= res_nocase$final_age[j] & res$sex == res_nocase$sex[j] & res$PTV_burden_group == res_nocase$PTV_burden_group[j]] + res_nocase$no_case[j]
      }
      res$PTV_burden_group = factor(res$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
      res %<>% 
        mutate(inc = 100000*case/(case2+no_case)) %>% 
        mutate(tumor = x, .before = "inc_time") %>% 
        mutate(PTVcutoff = co, .before = "inc_time")
      res$inc[is.nan(res$inc)] = 0

      res2 = expand.grid(inc_time_int = c("41_45", "46_50", "51_55", "56_60", "61_65", "66_70", "71_75", "76_80", "81_85", "86_90"), sex = unique(ukb_data$sex), PTV_burden_group = levels(tempdf_x$PTV_burden_group), stringsAsFactors = F) 
      res2$case = apply(res2, 1, function(y) {
        res_temp = res %>% filter(inc_time >= as.numeric(substr(y[1],1,2)), inc_time <= as.numeric(substr(y[1],4,5)), sex == as.numeric(y[2]), PTV_burden_group == y[3])
        sum(res_temp$case)
      })
      
      res2$case2 = apply(res2, 1, function(y) {
        res_temp = res %>% filter(inc_time >= as.numeric(substr(y[1],1,2)), inc_time <= as.numeric(substr(y[1],4,5)), sex == as.numeric(y[2]), PTV_burden_group == y[3])
        if(nrow(res_temp) == 0) {
          out = 0
        } else {
          out = res_temp$case2[nrow(res_temp)]
        }
        out
      })
      
      res2$no_case = apply(res2, 1, function(y) {
        res_temp = res %>% filter(inc_time >= as.numeric(substr(y[1],1,2)), inc_time <= as.numeric(substr(y[1],4,5)), sex == as.numeric(y[2]), PTV_burden_group == y[3])
        if(nrow(res_temp) == 0) {
          out = 0
        } else {
          out = res_temp$no_case[nrow(res_temp)]
        }
        out
      })
      res2$PTV_burden_group = factor(res2$PTV_burden_group, levels = levels(tempdf_x$PTV_burden_group))
      res2 %<>% 
        mutate(inc = 100000*case/(case2+no_case)) %>% 
        mutate(tumor = x, .before = "inc_time_int") %>% 
        mutate(PTVcutoff = co, .before = "inc_time_int")
      res2$inc_time_int = factor(res2$inc_time_int, levels = c("41_45", "46_50", "51_55", "56_60", "61_65", "66_70", "71_75", "76_80", "81_85", "86_90"))
      res2$inc_time_int_value = 1:(nrow(res2)/4)
      res2$inc[is.nan(res2$inc)] = 0
      w0 = wilcox.test(inc ~ PTV_burden_group, res2[res2$sex == 0,])$p.value
      w1 = wilcox.test(inc ~ PTV_burden_group, res2[res2$sex == 1,])$p.value
      ggarrange(plotlist = list(
        ggplot(res2[res2$sex == 0,], mapping = aes(x = inc_time_int_value, y = inc, color = PTV_burden_group)) + 
          geom_point() + 
          geom_smooth() +
          labs(title = paste0(x, " Women P = ", round(w0,4))) +
          scale_x_continuous(breaks = 1:(nrow(res2)/4), labels = c("41_45", "46_50", "51_55", "56_60", "61_65", "66_70", "71_75", "76_80", "81_85", "86_90")) +
          theme(axis.text.x = element_text(angle = 90)),
        ggplot(res2[res2$sex == 1,], mapping = aes(x = inc_time_int_value, y = inc, color = PTV_burden_group)) + 
          geom_point() + 
          geom_smooth() +
          labs(title = paste0(x, " Men P = ", round(w1,4))) +
          scale_x_continuous(breaks = 1:(nrow(res2)/4), labels = c("41_45", "46_50", "51_55", "56_60", "61_65", "66_70", "71_75", "76_80", "81_85", "86_90")) +
          theme(axis.text.x = element_text(angle = 90))
      ), ncol = 2, common.legend = T)
    })
    fig1 = ggarrange(plotlist = age_inc_plots, ncol = 5, nrow = 5, common.legend = T)
    ggsave(filename = paste0("plots/tumor_incidence_age_intervals/tumor_incidence_", ptv_burden_metrics[i], "_cutoff_", co ,".jpg"), plot = fig1, width = 100, height = 60, units = "cm", dpi = 300)
  })
}

```

