---
title: "incidence_from_attending"
author: "Bagi Laura"
date: "2024-01-19"
output: html_document
---


#Libraries
```{r}
library(ggpubr)
library(fastmatch)
library(readxl)
library(forcats)
library(survival)
library(survminer)
library(grid)
library(gridExtra)
library(data.table)
library(caret)
library(tidyverse)
library(future.apply)
library(magrittr)
library(cowplot)
options(dplyr.summarise.inform = F)

```

#Files
```{r}
load("C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/ukb_data_cancer_final")
load("C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/icd_mapping")
```

#Kaplen-Mayer
```{r}
plotlist_3 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
 tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, !(eid %in% tempdf_x$eid))
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$time_in_days = as.numeric(ifelse(!is.na(unique_tempdf$diag_date), unique_tempdf$diag_date - unique_tempdf$attending_1, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$attending_1))
  unique_tempdf = unique_tempdf[unique_tempdf$time_in_days >= 0, ,drop = F]
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  surv_obj = with(unique_tempdf, Surv(time_in_days, event = event_indicator))
  surv_model = survfit(surv_obj ~ 1)
  figure_2 = ggsurvplot(surv_model, data = unique_tempdf, title = paste0("Incidence Analysis (2) for ", x, "\n(N=", nrow(tempdf_x), ")"))
  figure_2$plot + ylim(1-(nrow(tempdf_x)/nrow(unique_tempdf)) - (nrow(tempdf_x)/nrow(unique_tempdf))*0.05, 1)
})
ggarrange(plotlist =  plotlist_3, ncol = 5, nrow = 5, common.legend = T)
ggsave(filename = "C:/Users/bagil/Desktop/ukb/plots/Survival/cancersurvival_issue_3.jpg", width = 60, height = 30, units = "cm", dpi = 300)

#with table
c_list_2 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, !(eid %in% tempdf_x$eid))
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$time_in_days = as.numeric(ifelse(!is.na(unique_tempdf$diag_date), unique_tempdf$diag_date - unique_tempdf$attending_1, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$attending_1))
  unique_tempdf = unique_tempdf[unique_tempdf$time_in_days >= 0, ,drop = F]
  unique_tempdf$cause_of_death_clean = substr(unique_tempdf$cause_of_death, 1, 3)
  unique_tempdf$p_cause_of_death = icd_mapping$cancer_type[match(unique_tempdf$cause_of_death_clean, icd_mapping$ICD10)]
  unique_tempdf$p_cause_of_death[is.na(unique_tempdf$p_cause_of_death)]="Not cancer"
  unique_tempdf$cancer_type[is.na(unique_tempdf$cancer_type)] = "No cancer"
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x & unique_tempdf$p_cause_of_death == x, 1, 0)
  surv_obj = with(unique_tempdf, Surv(time_in_days, event = event_indicator))
  surv_model = survfit(surv_obj ~ 1)
  figure_2 = ggsurvplot(surv_model, data = unique_tempdf, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), risk.table = T)
  figure_2$plot + ylim(1-(nrow(tempdf_x)/nrow(unique_tempdf)) - (nrow(tempdf_x)/nrow(unique_tempdf))*0.05, 1)
})
ggarrange(plotlist = c_list_2, nrow=5, ncol=5)
ggsave(filename = "C:/Users/bagil/Desktop/ukb/plots/Survival/cancer_issue3_table.jpg", width = 60, height = 30, units = "cm", dpi = 300)

```

#PTV group 3
```{r}

plotlist_PTV3a = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, !(eid %in% tempdf_x$eid))
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$time_in_days = as.numeric(ifelse(!is.na(unique_tempdf$diag_date), unique_tempdf$diag_date - unique_tempdf$attending_1, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$attending_1))
  unique_tempdf = unique_tempdf[unique_tempdf$time_in_days >= 0, ,drop = F]
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_3, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_3, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV3a, ncol = 5, nrow = 5)

ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_burden_g3_surv_a.jpg", width = 100, height = 50, units = "cm", dpi = 300)

```


```{r}
plotlist_PTV2a = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, !(eid %in% tempdf_x$eid))
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$time_in_days = as.numeric(ifelse(!is.na(unique_tempdf$diag_date), unique_tempdf$diag_date - unique_tempdf$attending_1, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$attending_1))
  unique_tempdf = unique_tempdf[unique_tempdf$time_in_days >= 0, ,drop = F]
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_2, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_2, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV2a, ncol = 5, nrow = 5)

ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_burden_g2_surv_a.jpg", width = 100, height = 50, units = "cm", dpi = 300)


```


```{r}
plotlist_PTV1a = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, !(eid %in% tempdf_x$eid))
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$time_in_days = as.numeric(ifelse(!is.na(unique_tempdf$diag_date), unique_tempdf$diag_date - unique_tempdf$attending_1, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$attending_1))
  unique_tempdf = unique_tempdf[unique_tempdf$time_in_days >= 0, ,drop = F]
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1a, ncol = 5, nrow = 5)

ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_burden_g1_surv_a.jpg", width = 100, height = 50, units = "cm", dpi = 300)


```


#Cos Model - sex
```{r}
cox_model_issue3_sex = sapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x & unique_tempdf$p_cause_of_death == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ sex, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  })

save(file = "C:/Users/bagil/Desktop/ukb/objects/coxmodel_issue3_sex", cox_model_issue3_sex)
```

#Cos Model - age
```{r}
cox_model_issue3_age_1 = sapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = arrange(tempdf_x, diag_date)
  tempdf_x$age = as.numeric(tempdf_x$diag_date - tempdf_x$birthdate)
  tempdf_no_x = subset(ukb_data_cancer, !(eid %in% tempdf_x$eid))
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_no_x$age = ifelse(tempdf_no_x$death == T, tempdf_no_x$date_of_death - tempdf_no_x$birthdate, max(unique_tempdf$date_of_death, na.rm = T) - tempdf_no_x$birthdate)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$age_in_years = floor(as.numeric(unique_tempdf$age)/365.25)
  #age_intervals = c(min(unique_tempdf$age_in_years), 19, 30, 40, 50, 60, 70, 80, max(unique_tempdf$age_in_years))
  #unique_tempdf$age_cat = cut(unique_tempdf$age_in_years, breaks = age_intervals, labels = F, include.lowest = T)
  unique_tempdf$time_in_days = as.numeric(ifelse(!is.na(unique_tempdf$diag_date), unique_tempdf$diag_date - unique_tempdf$attending_1, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$attending_1))
  unique_tempdf = unique_tempdf[unique_tempdf$time_in_days >= 0, ,drop = F]
  #unique_tempdf$cause_of_death_clean = substr(unique_tempdf$cause_of_death, 1, 3)
  #unique_tempdf$p_cause_of_death = icd_mapping$cancer_type[match(unique_tempdf$cause_of_death_clean, icd_mapping$ICD10)]
  #unique_tempdf$p_cause_of_death[is.na(unique_tempdf$p_cause_of_death)]="Not cancer"
  unique_tempdf$cancer_type[is.na(unique_tempdf$cancer_type)] = "No cancer"
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ age_in_years, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  })


#Age of first attending
cox_model_issue3_age_2 = sapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, !(eid %in% tempdf_x$eid))
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$age = as.numeric(ifelse(!is.na(unique_tempdf$attending_1), unique_tempdf$attending_1 - unique_tempdf$birthdate, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$birthdate))
  unique_tempdf$age_in_years = floor(as.numeric(unique_tempdf$age)/365.25)
  #age_intervals = c(0, 19, 30, 40, 50, 60, 70, 80, Inf)
  #unique_tempdf$age_cat = cut(unique_tempdf$age_in_years, breaks = age_intervals, labels = F)
  unique_tempdf$time_in_days = as.numeric(ifelse(!is.na(unique_tempdf$diag_date), unique_tempdf$diag_date - unique_tempdf$attending_1, max(unique_tempdf$date_of_death, na.rm = T) - unique_tempdf$attending_1))
  unique_tempdf = unique_tempdf[unique_tempdf$time_in_days >= 0, ,drop = F]
  #unique_tempdf$cause_of_death_clean = substr(unique_tempdf$cause_of_death, 1, 3)
  #unique_tempdf$p_cause_of_death = icd_mapping$cancer_type[match(unique_tempdf$cause_of_death_clean, icd_mapping$ICD10)]
  #unique_tempdf$p_cause_of_death[is.na(unique_tempdf$p_cause_of_death)]="Not cancer"
  unique_tempdf$cancer_type[is.na(unique_tempdf$cancer_type)] = "No cancer"
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ age_in_years, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  })

save(file = "C:/Users/bagil/Desktop/ukb/objects/coxmodel_issue3_age_1", cox_model_issue3_age_1)
save(file = "C:/Users/bagil/Desktop/ukb/objects/coxmodel_issue3_age_2", cox_model_issue3_age_2)

```

#Cos Model - ethnic
```{r}
cox_model_issue3_ethnic = sapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ ethnic_group, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  })

save(file = "C:/Users/bagil/Desktop/ukb/objects/coxmodel_issue3_ethnic", cox_model_issue3_ethnic)
```

#Cos Model - skin
```{r}

cox_model_issue3_skin = sapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ s.color, data = unique_tempdf)
  figure = ggsurvplot(survfit(res.cox), data = unique_tempdf)
  figure$plot + ylim(1-(nrow(tempdf_x)/nrow(unique_tempdf)) - (nrow(tempdf_x)/nrow(unique_tempdf))*0.05, 1)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  })

save(file ="C:/Users/bagil/Desktop/ukb/objects/coxmodel_issue3_skin", cox_model_issue3_skin)
```

#PTV cox group 1
```{r}

coxregfun_a_g1 = function(x) {
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden_group, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value,
    group = c("group1")
  )
}

result_list = lapply(unique(icd_mapping$cancer_type), coxregfun_a_g1)


cancer_new_PTV_g1_a_df = do.call(rbind, result_list)
save(cancer_new_PTV_g1_a_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_g1_attending")


```

#PTV cox grup 2
```{r}
coxregfun_a_g2 = function(x) {
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden_group_2, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value,
    group = c("group2")
  )
}

result_list_2 = lapply(unique(icd_mapping$cancer_type), coxregfun_a_g2)


cancer_new_PTV_g2_a_df = do.call(rbind, result_list_2)
save(cancer_new_PTV_g2_a_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_g2_attending")


```

#PTV cox group 3
```{r}

coxregfun_a_g3 = function(x) {
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden_group_3, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value,
    group = c("group3")
  )
}

result_list_3 = lapply(unique(icd_mapping$cancer_type), coxregfun_a_g3)


cancer_new_PTV_g3_a_df = do.call(rbind, result_list_3)
save(cancer_new_PTV_g3_a_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_g3_attending")


```

#PTV cox
```{r}

coxregfun_a = function(x) {
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden, data = unique_tempdf)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value)
}

result_list_a = lapply(unique(icd_mapping$cancer_type), coxregfun_a)


cancer_new_PTV_a_df = do.call(rbind, result_list_a)
save(cancer_new_PTV_a_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_attending")


```


#PTV gruop 1 survival
```{r}

ukb_data_cancer$cancer_type[is.na(ukb_data_cancer$cancer_type)] = "No_cancer"

plotlist_PTV1 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$PTV_burden_group_5per = factor(unique_tempdf$PTV_burden_group_5per, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_5per, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_5per, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/new_PTV_burden_5per_attending.png", width = 100, height = 50, units = "cm", dpi = 300)


```

#PTV gruop 2 survival
```{r}

plotlist_PTV2 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$PTV_burden_group_10per = factor(unique_tempdf$PTV_burden_group_10per, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_10per, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_10per, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV2, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/new_PTV_burden_10per_attending.png", width = 100, height = 50, units = "cm", dpi = 300)

```

#PTV gruop 3 survival
```{r}
plotlist_PTV3 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$PTV_burden_grou_3 = factor(unique_tempdf$PTV_burden_group_3, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_3, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_3, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV3, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/new_PTV_burden_g3_attending.png", width = 100, height = 50, units = "cm", dpi = 300)


```

#Cancer all incidence
```{r}

#table creating
temp_df_3 = data.frame(eid = ukb_data_cancer$eid,
                       birthdate = ukb_data_cancer$birthdate,
                       attending = ukb_data_cancer$attending_1,
                       date_of_death = ukb_data_cancer$date_of_death,
                       diag_date = ukb_data_cancer$diag_date,
                       death = ukb_data_cancer$death,
                       cancer_type = ukb_data_cancer$cancer_type,
                       PTV_burden_group = ukb_data_cancer$PTV_burden_group,
                       PTV_burden_group_2 = ukb_data_cancer$PTV_burden_group_2,
                       PTV_burden_group_3 = ukb_data_cancer$PTV_burden_group_3)

temp_df_3$cancer_type[is.na(temp_df_3$cancer_type)] = "No_cancer"

temp_df_3 <- temp_df_3 %>%
  mutate(has_cancer = if_else(cancer_type != "No_cancer", TRUE, FALSE))

temp_df_3 %<>%
  group_by(eid) %>%
  summarize(
    birthdate = first(birthdate),
    attending = first(attending),
    date_of_death = first(date_of_death),
    diag_date = first(diag_date),
    death = first(death),
    has_cancer = first(has_cancer),
    PTV_burden_group = first(PTV_burden_group),
    PTV_burden_group_2 = first(PTV_burden_group_2),
    PTV_burden_group_3 = first(PTV_burden_group_3))

temp_df_3 %<>% 
  mutate(time_in_days = case_when(
      has_cancer == T & death == F ~ diag_date - attending,
      has_cancer == T & death == T ~ date_of_death - attending,
      death == T & has_cancer == F ~ date_of_death - attending,
      death == F & has_cancer == F ~ max(date_of_death, na.rm = T) - attending
    ))

#Surv. plots
fit_7 = survfit(Surv(time_in_days, event = has_cancer) ~ PTV_burden_group, data = temp_df_3)
PTV_g1_cancer_surv_a = ggsurvplot(fit_7, risk.table = T, pval = T, tables.theme = clean_theme(), title = paste0("Incidence Analysis for Tumors from attending (PTV burden group 1)"))
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_g1_tumor_incidence_attending.jpg", plot = PTV_g1_cancer_surv_a$plot, width = 60, height = 30, units = "cm", dpi = 300)

fit_8 = survfit(Surv(time_in_days, event = has_cancer) ~ PTV_burden_group_2, data = temp_df_3)
PTV_g2_cancer_surv_a = ggsurvplot(fit_8, risk.table = T, pval = T, tables.theme = clean_theme(), title = paste0("Incidence Analysis for Tumors from attending (PTV burden group 2)"))
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_g2_tumor_incidence.jpg", plot = PTV_g2_cancer_surv_a$plot, width = 60, height = 30, units = "cm", dpi = 300)

fit_9 = survfit(Surv(time_in_days, event = has_cancer) ~ PTV_burden_group_3, data = temp_df_3)
PTV_g3_cancer_surv_a = ggsurvplot(fit_9, risk.table = T, pval = T, tables.theme = clean_theme(), title = paste0("Incidence Analysis for Tumors from attending (PTV burden group 3)"))
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_g3_tumor_incidence.jpg", plot = PTV_g3_cancer_surv_a$plot, width = 60, height = 30, units = "cm", dpi = 300)


#plots and risk table combine
PTV_plot_list_inc_a = list(PTV_g1_cancer_surv_a$plot, PTV_g2_cancer_surv_a$plot, PTV_g3_cancer_surv_a$plot)
PTV_risk_list_inc_a = list(PTV_g1_cancer_surv_a$table, PTV_g2_cancer_surv_a$table, PTV_g3_cancer_surv_a$table)
PTV_plot_list_combined_inc_a = c(PTV_plot_list_inc_a, PTV_risk_list_inc_a)

grid.arrange(grobs = PTV_plot_list_combined_inc_a, ncol = 3)
plot_grid(plotlist = PTV_plot_list_combined_inc_a, ncol = 3)
ggsave(filename ="C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_cancer_incidence_attending.jpg", width = 100, height = 50, units = "cm", dpi = 300 )


```

#Forest plot
```{r}

tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]
i = 0

PTV_tumor_attending_forest_plot = lapply(X = tumors, FUN = function(x){
  i = i + 1
  .GlobalEnv$i = i
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  #icd_code = icd_mapping$ICD10[icd_mapping$cancer_type == x]
  #tempdf_x_dead = tempdf_x %>% filter(death == T, cause_of_death == icd_code)
  #tempdf_x_alive = tempdf_x %>% filter(death == F)
  #tempdf_x = rbind(tempdf_x_alive, tempdf_x_dead)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    ))
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$sex = factor(unique_tempdf$sex)
  l = tempdf_x %>% 
    add_count(ethnic_group, name = "freq") %>% 
    filter(freq > nrow(tempdf_x)/50) %>% 
    arrange(desc(freq)) %>% 
    pull(ethnic_group) %>% 
    unique()
  if (length(l) == 1 & length(unique(tempdf_x$sex)) == 1){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  } else if(length(l) == 1 & length(unique(tempdf_x$sex)) == 2){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  } else if(length(l) > 1 & length(unique(tempdf_x$sex)) == 1){
    unique_tempdf$ethnic_group = factor(unique_tempdf$ethnic_group, levels = c("British", sort(setdiff(l, "British"))))
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + ethnic_group + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  } else {
    unique_tempdf$ethnic_group = factor(unique_tempdf$ethnic_group, levels = c("British", sort(setdiff(l, "British"))))
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + ethnic_group + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  }
  ggforest(hr_model, data = as.data.frame(unique_tempdf), main = paste0("Hazard ratio of ", x))
})

PTV_attending_cox_forest = ggarrange(plotlist = PTV_tumor_attending_forest_plot, ncol = 9, nrow = 3)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/tumor_forest_a.png", plot = PTV_attending_cox_forest, width = 120, height = 60, units = "cm", limitsize = F, dpi = 300)


```

#Forest plot tumor categories
```{r}

#load("C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/ukb_data_cancer_final")
hist_cat = unique(ukb_data_cancer$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]
i = 0

PTV_tumor_attending_forest_plot = lapply(X = hist_cat, FUN = function(x){
  i = i + 1
  .GlobalEnv$i = i
  tempdf_x = subset(ukb_data_cancer, hist_tumor_cat == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  #icd_code = icd_mapping$ICD10[icd_mapping$cancer_type == x]
  #tempdf_x_dead = tempdf_x %>% filter(death == T, cause_of_death == icd_code)
  #tempdf_x_alive = tempdf_x %>% filter(death == F)
  #tempdf_x = rbind(tempdf_x_alive, tempdf_x_dead)
  tempdf_no_x = subset(ukb_data_cancer, hist_tumor_cat == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      hist_tumor_cat == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    ))
  unique_tempdf$event_indicator = ifelse(unique_tempdf$hist_tumor_cat == x, 1, 0)
  unique_tempdf$sex = factor(unique_tempdf$sex)
  l = tempdf_x %>% 
    add_count(ethnic_group, name = "freq") %>% 
    filter(freq > nrow(tempdf_x)/50) %>% 
    arrange(desc(freq)) %>% 
    pull(ethnic_group) %>% 
    unique()
  if (length(l) == 1 & length(unique(tempdf_x$sex)) == 1){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  } else if(length(l) == 1 & length(unique(tempdf_x$sex)) == 2){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  } else if(length(l) > 1 & length(unique(tempdf_x$sex)) == 1){
    unique_tempdf$ethnic_group = factor(unique_tempdf$ethnic_group, levels = c("British", sort(setdiff(l, "British"))))
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + ethnic_group + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  } else {
    unique_tempdf$ethnic_group = factor(unique_tempdf$ethnic_group, levels = c("British", sort(setdiff(l, "British"))))
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + ethnic_group + gpca_1 + gpca_2 + gpca_3 + gpca_4 + gpca_5 + gpca_6 + gpca_7 + gpca_8 + gpca_9 + gpca_10, data = unique_tempdf)
  }
  ggforest(hr_model, data = as.data.frame(unique_tempdf), main = paste0("Hazard ratio of ", x))
})

PTV_attending_cox_forest = ggarrange(plotlist = PTV_tumor_attending_forest_plot, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/tumor_cat_forest_a.png", plot = PTV_attending_cox_forest, width = 120, height = 50, units = "cm", limitsize = F, dpi = 300)


```


#Surv analysis
##Tumor categories universal 1
```{r}

#load("C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/ukb_data_cancer_final")
hist_cat = unique(ukb_data_cancer$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

plotlist_PTV1 = lapply(X = hist_cat, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, hist_tumor_cat == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, hist_tumor_cat == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      hist_tumor_cat == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$hist_tumor_cat == x, 1, 0)
  #unique_tempdf$PTV_universal_1 = factor(unique_tempdf$PTV_universal_1, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_universal_1, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_universal_1, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_universal_1_attending.png", width = 160, height = 70, units = "cm", dpi = 300, limitsize = F)


```

##Tumor categories universal 2
```{r}

hist_cat = unique(ukb_data_cancer$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

plotlist_PTV1 = lapply(X = hist_cat, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, hist_tumor_cat == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, hist_tumor_cat == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      hist_tumor_cat == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$hist_tumor_cat == x, 1, 0)
  #unique_tempdf$PTV_universal_2 = factor(unique_tempdf$PTV_universal_2, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_universal_2, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_universal_2, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_universal_2_attending.png", width = 160, height = 70, units = "cm", dpi = 300, limitsize = F)

```

##Surv analysis for tumor categories cancer
```{r}
hist_cat = unique(ukb_data_cancer$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

plotlist_PTV1 = lapply(X = hist_cat, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, hist_tumor_cat == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, hist_tumor_cat == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      hist_tumor_cat == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$hist_tumor_cat == x, 1, 0)
  unique_tempdf$PTV_cancer = factor(unique_tempdf$PTV_cancer, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_cancer_attending.png", width = 170, height = 70, units = "cm", dpi = 300, limitsize = F)


```

##Tumor categories specific 1
```{r}

hist_cat = unique(ukb_data_cancer$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

plotlist_PTV1 = lapply(X = hist_cat, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, hist_tumor_cat == x)
  tempdf_x$PTV_hist_type_1 = as.character(tempdf_x$PTV_hist_type_1)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  l = sort(unique(tempdf_x$PTV_hist_type_1))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  tempdf_no_x = subset(ukb_data_cancer, hist_tumor_cat == "No_cancer")
  tempdf_no_x$PTV_hist_type_1 = as.character(tempdf_no_x$PTV_hist_type_1)
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_no_x$PTV_hist_type_1 = cut(tempdf_no_x$PTV_burden, breaks = cp, include.lowest = T, right = T)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$PTV_hist_type_1 = cut(unique_tempdf$PTV_burden, breaks = cp, include.lowest = T, right = T)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      hist_tumor_cat == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$hist_tumor_cat == x, 1, 0)
  unique_tempdf$PTV_hist_type_1 = factor(unique_tempdf$PTV_hist_type_1, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_1, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_1, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_tumor_cat_spec_1_attending.png", width = 170, height = 70, units = "cm", dpi = 300, limitsize = F)


```

##Tumor categories specific 2
```{r}
hist_cat = unique(ukb_data_cancer$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

plotlist_PTV1 = lapply(X = hist_cat, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, hist_tumor_cat == x)
  tempdf_x$PTV_hist_type_2 = as.character(tempdf_x$PTV_hist_type_2)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  l = sort(unique(tempdf_x$PTV_hist_type_2))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  tempdf_no_x = subset(ukb_data_cancer, hist_tumor_cat == "No_cancer")
  tempdf_no_x$PTV_hist_type_2 = as.character(tempdf_no_x$PTV_hist_type_2)
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_no_x$PTV_hist_type_2 = cut(tempdf_no_x$PTV_burden, breaks = cp, include.lowest = T, right = T)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$PTV_hist_type_2 = cut(unique_tempdf$PTV_burden, breaks = cp, include.lowest = T, right = T)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      hist_tumor_cat == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$hist_tumor_cat == x, 1, 0)
  #unique_tempdf$PTV_hist_type_2 = factor(unique_tempdf$PTV_hist_type_2, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_2, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_2, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_tumor_cat_spec_2_attending.png", width = 170, height = 70, units = "cm", dpi = 300, limitsize = F)


```

##Hist cat 3
```{r}

hist_cat = unique(ukb_data_cancer$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

plotlist_PTV1 = lapply(X = hist_cat, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, hist_tumor_cat == x)
  tempdf_x$PTV_hist_type_3 = as.character(tempdf_x$PTV_hist_type_3)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  l = sort(unique(tempdf_x$PTV_hist_type_3))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  tempdf_no_x = subset(ukb_data_cancer, hist_tumor_cat == "No_cancer")
  tempdf_no_x$PTV_hist_type_3 = as.character(tempdf_no_x$PTV_hist_type_3)
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_no_x$PTV_hist_type_3 = cut(tempdf_no_x$PTV_burden, breaks = cp, include.lowest = T, right = T)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$PTV_hist_type_3 = cut(unique_tempdf$PTV_burden, breaks = cp, include.lowest = T, right = T)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      hist_tumor_cat == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$hist_tumor_cat == x, 1, 0)
  #unique_tempdf$PTV_hist_type_3 = factor(unique_tempdf$PTV_hist_type_3, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_3, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_3, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/PTV_tumor_cat_spec_3_attending.png", width = 170, height = 70, units = "cm", dpi = 300, limitsize = F)

```


##Cancer types tumor spec
```{r}

tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

plotlist_PTV1 = lapply(X = tumors, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x$PTV_cancer_type_1 = as.character(tempdf_x$PTV_cancer_type_1)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  l = sort(unique(tempdf_x$PTV_cancer_type_1))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x$PTV_cancer_type_1 = as.character(tempdf_no_x$PTV_cancer_type_1)
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_no_x$PTV_cancer_type_1 = cut(tempdf_no_x$PTV_burden, breaks = cp, include.lowest = T, right = T)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$PTV_cancer_type_1 = cut(unique_tempdf$PTV_burden, breaks = cp, include.lowest = T, right = T)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$PTV_cancer_type_1 = factor(unique_tempdf$PTV_cancer_type_1, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_1, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_1, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/cancer_type_1_PTV_attending_cancer.png", width = 150, height = 80, units = "cm", dpi = 300, limitsize = F)



```


```{r}

tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

plotlist_PTV1 = lapply(X = tumors, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x$PTV_cancer_type_2 = as.character(tempdf_x$PTV_cancer_type_2)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  l = sort(unique(tempdf_x$PTV_cancer_type_2))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x$PTV_cancer_type_2 = as.character(tempdf_no_x$PTV_cancer_type_2)
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_no_x$PTV_cancer_type_2 = cut(tempdf_no_x$PTV_burden, breaks = cp, include.lowest = T, right = T)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$PTV_cancer_type_2 = cut(unique_tempdf$PTV_burden, breaks = cp, include.lowest = T, right = T)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  #unique_tempdf$PTV_cancer_type_2 = factor(unique_tempdf$PTV_cancer_type_2, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_2, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_2, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/cancer_type_2_PTV_attending_cancer.png", width = 150, height = 80, units = "cm", dpi = 300, limitsize = F)

```

##CAncer type 3
```{r}

tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

plotlist_PTV1 = lapply(X = tumors, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x$PTV_cancer_type_3 = as.character(tempdf_x$PTV_cancer_type_3)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  l = sort(unique(tempdf_x$PTV_cancer_type_3))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x$PTV_cancer_type_3 = as.character(tempdf_no_x$PTV_cancer_type_3)
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_no_x$PTV_cancer_type_3 = cut(tempdf_no_x$PTV_burden, breaks = cp, include.lowest = T, right = T)
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf$PTV_cancer_type_3 = cut(unique_tempdf$PTV_burden, breaks = cp, include.lowest = T, right = T)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  #unique_tempdf$PTV_cancer_type_3 = factor(unique_tempdf$PTV_cancer_type_3, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_3, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_3, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/cancer_type_3_PTV_attending_cancer.png", width = 150, height = 80, units = "cm", dpi = 300, limitsize = F)

```


##Cancer types universal 1
```{r}

tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

plotlist_PTV1 = lapply(X = tumors, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$PTV_universal_1 = factor(unique_tempdf$PTV_universal_1, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_universal_1, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_universal_1, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/cancer_type_PTV_universal_1_attending.png", width = 150, height = 80, units = "cm", dpi = 300, limitsize = F)

```

##Cancer types universal 2
```{r}

tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

plotlist_PTV1 = lapply(X = tumors, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$PTV_universal_2 = factor(unique_tempdf$PTV_universal_2, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_universal_2, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_universal_2, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/cancer_type_PTV_universal_2_attending.png", width = 150, height = 80, units = "cm", dpi = 300, limitsize = F)

```


#Survival analysis for cancer types cancer
```{r}
tumors = unique(ukb_data_cancer$cancer_type)
tumors = tumors[tumors != "No_cancer"]

plotlist_PTV1 = lapply(X = tumors, FUN = function(x){
  tempdf_x = subset(ukb_data_cancer, cancer_type == x)
  tempdf_x = dplyr::arrange(tempdf_x, diag_date)
  tempdf_no_x = subset(ukb_data_cancer, cancer_type == "No_cancer")
  tempdf_no_x = tempdf_no_x[!duplicated(tempdf_no_x$eid), ]
  tempdf_x = tempdf_x[!duplicated(tempdf_x$eid), ]
  unique_tempdf = rbind(tempdf_x, tempdf_no_x)
  unique_tempdf %<>% 
    mutate(time_in_days = case_when(
      cancer_type == x ~ diag_date - attending_1,
      death == T & is.na(diag_date) ~ date_of_death - attending_1,
      death == F & is.na(diag_date) ~ max(date_of_death, na.rm = T) - attending_1,
    )) %>% 
    filter(time_in_days >= 0)
  unique_tempdf$event_indicator = ifelse(unique_tempdf$cancer_type == x, 1, 0)
  unique_tempdf$PTV_cancer = factor(unique_tempdf$PTV_cancer, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer, data = unique_tempdf))
  args = list(fit = fit, title = paste0("Incidence Analysis for ", x, "\n(N=", nrow(tempdf_x), ")"), ylim = c(1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05, 1), pval = T, pval.coord = c(1, 1 - (nrow(tempdf_x) / nrow(unique_tempdf)) - (nrow(tempdf_x) / nrow(unique_tempdf)) * 0.05),legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer, data = unique_tempdf), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  plotlist_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Incidence/Incidence_from_attending/cancer_type_PTV_cancer_attending.png", width = 150, height = 80, units = "cm", dpi = 300, limitsize = F)



```

