---
title: "PTV_burden"
author: "Balazs Koncz"
date: "2024-09-17"
output: html_document
---

#Setup

```{r}
setwd("/media/balazs/WorkL/balazs/Work/UKBiobank/")
Packages <- c("tidyverse", "data.table", "magrittr", "fastmatch", "ggplot2", "pbapply", "clusterProfiler", "org.Hs.eg.db", "Rfast", "future.apply")
lapply(Packages, require, character.only = TRUE)
rm(Packages)
options(dplyr.summarise.inform = FALSE)
```

#GO terms

```{r}
go_terms = fread("Objects/go_terms.txt") #https://amigo.geneontology.org/amigo/search/bioentity
go_terms %<>% filter(grepl("UniProt", V1))
go_terms$V1 = gsub("UniProtKB:", "", go_terms$V1)
go_terms$V10 = gsub("PANTHER:", "", go_terms$V10)
go_terms = go_terms[,c(2,1,6,7,5,10,11)]
colnames(go_terms) = c("Gene", "UniProtKB", "Protein", "GO_terms", "GO_descriptions", "PANTHER_family", "PANTHER_description")

goterm_tbl = apply(go_terms, 1, function(x) {
  data.frame(Gene = x[1], 
             UniProtKB = x[2], 
             Protein = x[3], 
             GO_term = strsplit(x[4], "\\|")[[1]],
             GO_description = strsplit(x[5], "\\|")[[1]],
             PANTHER_family = x[6],
             PANTHER_description = x[7])
})
table(sapply(goterm_tbl, ncol))
goterm_tbl %<>% bind_rows()

save(goterm_tbl, file = "Objects/goterm_tbl")
```

#PTVburden_without_MAF-001.tsv

```{r}
#Load PTV burden data
ptvb = fread("Objects/PTVburden_without_MAF-001/PTVburden_without_MAF-001.tsv")

genes_all = data.frame(ENSEMBL = sort(unique(unlist(strsplit(ptvb$Genes, ","), use.names = F))))
gene_symbols = bitr(geneID = genes_all$ENSEMBL, fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
genes_all$SYMBOL = gene_symbols$SYMBOL[fmatch(genes_all$ENSEMBL, gene_symbols$ENSEMBL)]
rm(gene_symbols)

```

##Immune-related genes (using GO terms)

```{r}
load("Objects/goterm_tbl")
genes_immune = genes_all %>% filter(SYMBOL %in% goterm_tbl$Gene[grepl("immune", goterm_tbl$GO_description)]) #1120

```

##Tumor suppressor genes

```{r}
tsg = fread("Objects/tumor_supressor_genes_bioinfo_uth.txt")
genes_tsg = genes_all %>% filter(SYMBOL %in% tsg$GeneSymbol) #908

```

##pLI genes

```{r}
pli = fread("Objects/LoF_intolerant_genes.tsv")
genes_pli = genes_all %>% filter(SYMBOL %in% pli$gene) #2805

```

##Calculate specific PTV burden scores

```{r}
temp = pblapply(ptvb$Genes, function(x) {
  g = unique(strsplit(x, ",")[[1]])
  c(sum(g %fin% genes_immune$ENSEMBL),
    sum(g %fin% genes_tsg$ENSEMBL),
    sum(g %fin% genes_pli$ENSEMBL))
})
temp2 = do.call(rbind, temp)
colnames(temp2) = c("PTVim", "PTVtsg", "PTVpli")
ptvb = cbind(ptvb, temp2)

ptvb = as.data.frame(ptvb[,c(1,2,4,6,7,8)])
colnames(ptvb) = c("eid", "PTV", "Genes_PTV", "PTVim", "PTVtsg", "PTVpli")

ggplot(ptvb, aes(x = PTVim)) + geom_histogram()
ggplot(ptvb, aes(x = PTVtsg)) + geom_histogram()
quantile(ptvb$PTVtsg, probs = seq(0,1,0.01))
quantile(ptvb$PTVim, probs = seq(0,1,0.01))
quantile(ptvb$PTVpli, probs = seq(0,1,0.01))

rm(genes_all, genes_immune, genes_pli, genes_tsg, temp, temp2)
```

#MAF10_4_all_retained_variants_PTVBurden_final_Shetscores.tsv
##Extract genes

```{r}
#Load PTV burden data
ptvb_wMAF = fread("Objects/PTV_2024_05/MAF10_4_all_retained_variants_PTVBurden_final_Shetscores.tsv")

genes_all = data.frame(ENSEMBL = sort(unique(unlist(strsplit(ptvb_wMAF$Genes, ","), use.names = F))))
gene_symbols = bitr(geneID = genes_all$ENSEMBL, fromType = "ENSEMBL", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
genes_all$SYMBOL = gene_symbols$SYMBOL[fmatch(genes_all$ENSEMBL, gene_symbols$ENSEMBL)]
rm(gene_symbols)

```

##Immune-related genes (using GO terms)

```{r}
genes_immune = genes_all %>% filter(SYMBOL %in% goterm_tbl$Gene[grepl("immune", goterm_tbl$GO_description)]) #1233
```

##Tumor suppressor genes

```{r}
genes_tsg = genes_all %>% filter(SYMBOL %in% tsg$GeneSymbol) #953
```

##pLI genes

```{r}
genes_pli = genes_all %>% filter(SYMBOL %in% pli$gene) #2866
```

##Calculate specific PTV burden scores

```{r}
temp = pblapply(ptvb_wMAF$Genes, function(x) {
  g = strsplit(x, ",")[[1]]
  c(sum(g %fin% genes_immune$ENSEMBL),
    sum(g %fin% genes_tsg$ENSEMBL),
    sum(g %fin% genes_pli$ENSEMBL))
})
temp2 = do.call(rbind, temp)
colnames(temp2) = c("PTVim_MAF", "PTVtsg_MAF", "PTVpli_MAF")
ptvb_wMAF = cbind(ptvb_wMAF, temp2)

ptvb_wMAF = as.data.frame(ptvb_wMAF[,c(1,2,4,7,8,9)])
colnames(ptvb_wMAF) = c("eid", "PTV_MAF", "Genes_PTV_MAF", "PTVim_MAF", "PTVtsg_MAF", "PTVpli_MAF")
```

#JOIN

```{r}
ptvb_data = ptvb %>% full_join(ptvb_wMAF, by = "eid") 
Table(ptvb$eid %in% ptvb_data$eid)
Table(ptvb_wMAF$eid %in% ptvb_data$eid)

saveRDS(ptvb_data, file = "Objects/PTV_data.rds")

rm(genes_all, genes_immune, genes_pli, genes_tsg, goterm_tbl, pli, ptvb, ptvb_wMAF, temp, temp2, tsg)
```


#Correlations

```{r}
ptvb_data = readRDS("Objects/PTV_data.rds")
ComplexHeatmap::Heatmap(cor(ptvb_data[,c(2,4:7,9:11)], method = "spearman"), cluster_rows = F, cluster_columns = F)

Table(ptvb_data[,7])
Table(ptvb_data[,9])
Table(ptvb_data[,10])
Table(ptvb_data[,11])

```

#Groups

```{r}
ptvb_data = readRDS("Objects/PTV_data.rds")
ptvb_data %>% pull(eid) %>% unique() %>% length()

#PTV
ptvb_data$PTV_g90 = cut(ptvb_data$PTV, breaks = c(min(ptvb_data$PTV, na.rm = T), quantile(ptvb_data$PTV, .9, na.rm = T), max(ptvb_data$PTV, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTV_g90)

ptvb_data$PTV_g95 = cut(ptvb_data$PTV, breaks = c(min(ptvb_data$PTV, na.rm = T), quantile(ptvb_data$PTV, .95, na.rm = T), max(ptvb_data$PTV, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTV_g95)

#PTVim
ptvb_data$PTVim_g90 = cut(ptvb_data$PTVim, breaks = c(min(ptvb_data$PTVim, na.rm = T), quantile(ptvb_data$PTVim, .9, na.rm = T), max(ptvb_data$PTVim, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTVim_g90)

ptvb_data$PTVim_g95 = cut(ptvb_data$PTVim, breaks = c(min(ptvb_data$PTVim, na.rm = T), quantile(ptvb_data$PTVim, .95, na.rm = T), max(ptvb_data$PTVim, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTVim_g95)

#PTVtsg
ptvb_data$PTVtsg_g90 = cut(ptvb_data$PTVtsg, breaks = c(min(ptvb_data$PTVtsg, na.rm = T), quantile(ptvb_data$PTVtsg, .9, na.rm = T), max(ptvb_data$PTVtsg, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTVtsg_g90)

ptvb_data$PTVtsg_g95 = cut(ptvb_data$PTVtsg, breaks = c(min(ptvb_data$PTVtsg, na.rm = T), quantile(ptvb_data$PTVtsg, .95, na.rm = T), max(ptvb_data$PTVtsg, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTVtsg_g95)

#PTVpli
ptvb_data$PTVpli_g90 = cut(ptvb_data$PTVpli, breaks = c(min(ptvb_data$PTVpli, na.rm = T), quantile(ptvb_data$PTVpli, .9, na.rm = T), max(ptvb_data$PTVpli, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTVpli_g90)

# ptvb_data$PTVpli_g95 = cut(ptvb_data$PTVpli, breaks = c(min(ptvb_data$PTVpli, na.rm = T), quantile(ptvb_data$PTVpli, .95, na.rm = T), max(ptvb_data$PTVpli, na.rm = T)), include.lowest = T)
# Table(ptvb_data$PTVpli_g95)
# ptvb_data$PTVpli_g95 = NULL

#PTV_MAF
ptvb_data$PTV_MAF_g90 = cut(ptvb_data$PTV_MAF, breaks = c(min(ptvb_data$PTV_MAF, na.rm = T), quantile(ptvb_data$PTV_MAF, .9, na.rm = T), max(ptvb_data$PTV_MAF, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTV_MAF_g90)

ptvb_data$PTV_MAF_g95 = cut(ptvb_data$PTV_MAF, breaks = c(min(ptvb_data$PTV_MAF, na.rm = T), quantile(ptvb_data$PTV_MAF, .95, na.rm = T), max(ptvb_data$PTV_MAF, na.rm = T)), include.lowest = T)
Table(ptvb_data$PTV_MAF_g95)

#PTVim_MAF
ptvb_data$PTVim_MAF_g = cut(ptvb_data$PTVim_MAF, breaks = c(min(ptvb_data$PTVim_MAF, na.rm = T), 1, max(ptvb_data$PTVim_MAF, na.rm = T)), include.lowest = T, right = F)
Table(ptvb_data$PTVim_MAF_g)

#PTVtsg_MAF
ptvb_data$PTVtsg_MAF_g = cut(ptvb_data$PTVtsg_MAF, breaks = c(min(ptvb_data$PTVtsg_MAF, na.rm = T), 1, max(ptvb_data$PTVtsg_MAF, na.rm = T)), include.lowest = T, right = F)
Table(ptvb_data$PTVtsg_MAF_g)

#PTVpli_MAF
ptvb_data$PTVpli_MAF_g = cut(ptvb_data$PTVpli_MAF, breaks = c(min(ptvb_data$PTVpli_MAF, na.rm = T), 1, max(ptvb_data$PTVpli_MAF, na.rm = T)), include.lowest = T, right = F)
Table(ptvb_data$PTVpli_MAF_g)

```

#Add UKB data

```{r}
load("Objects/ukb_data_cancer_final")
sum(Table(ukb_data_cancer$cancer_type))

ukb_data = ukb_data_cancer[,c(1:23,45:84)]
ukb_data %<>% left_join(ptvb_data, by = "eid")

# #GPCA
# gpca = fread("Objects/ukb_genetic_pca_40.tsv")
# gpca %<>% dplyr::select(eid, UKB_column_name, value) %>% pivot_wider(id_cols = eid, names_from = UKB_column_name, values_from = value)
# colnames(gpca) = gsub("22009-0.", "gpca", colnames(gpca))
# gpca$eid = as.character(gpca$eid)
# ukb_data %<>% left_join(gpca, by = "eid")

saveRDS(ukb_data, file = "Objects/ukb_data.rds")
```

#MAF from UKB

```{r}
#Load PTV burden data
ptvb = fread("PTVvars/PTVburden_without_MAF-001/PTVburden_without_MAF-001.tsv")
plan(multisession(workers = 4))
ptv_freq = Table(unlist(future_lapply(ptvb$Variants, function(x) strsplit(x, ",")[[1]]), use.names = F))
m = round(nrow(ptvb)/10000)
gc()

ukb_data = readRDS("Objects/ukb_data.rds")
ukb_data$cancer_type[ukb_data$cancer_type == "Brain/CNS"] = "Brain"
ukb_data$cancer_type[ukb_data$cancer_type == "Hodgkin lymphoma"] = "HodgkinLymphoma"
ukb_data$cancer_type[ukb_data$cancer_type == "Multiple myeloma"] = "MultipleMyeloma"
ukb_data$cancer_type[ukb_data$cancer_type == "Non-Hodgkin lymphoma"] = "NonHodgkinLymphoma"
ukb_data$sex = factor(ukb_data$sex, levels = c(0,1), labels = c("female", "male"))
```

