---
title: "Survival_analysis_cancer"
author: "Bagi Laura"
date: "2024-03-18"
output: html_document
---

#Libraries
```{r}
library(ggpubr)
library(fastmatch)
library(readxl)
library(forcats)
library(survival)
library(survminer)
library(grid)
library(gridExtra)
library(data.table)
library(caret)
library(tidyverse)
library(future.apply)
library(magrittr)
library(cowplot)
options(dplyr.summarise.inform = F)

```

#Files
```{r}
load("C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_data_final")
load("C:/Users/bagil/Desktop/ukb/objects/icd_mapping")
load("C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/ukb_data_cancer_final")

```


#Survival analysis for PTV burden group 5%
```{r}
cancer_surv_PTV1 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group_5per = factor(temp_can$PTV_burden_group_5per, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_5per,data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_5per, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/new_PTV_burden_5per_surv_cancer.png", width = 100, height = 50, units = "cm", dpi = 300)


```

#Survival analysis for PTV burden group 10%
```{r}

cancer_surv_PTV2 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group_10per = factor(temp_can$PTV_burden_group_10per, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_10per, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_10per, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  cancer_surv_PTV2, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/new_PTV_burden_10per_surv_cancer.png", width = 100, height = 50, units = "cm", dpi = 300)

```

#Survival analysis for PTV burden group 15%
```{r}

cancer_surv_PTV3 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group_15per = factor(temp_can$PTV_burden_group_15per, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_15per, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_15per, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  cancer_surv_PTV3, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/new_PTV_burden_15per_surv_cancer.png", width = 100, height = 50, units = "cm", dpi = 300)


```

#Survival analysis for PTV burden group 20%
```{r}

cancer_surv_PTV4 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group_20per = factor(temp_can$PTV_burden_group_20per, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_20per, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_20per, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})

ggarrange(plotlist =  cancer_surv_PTV4, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/new_PTV_burden_20per_surv_cancer.png", width = 100, height = 50, units = "cm", dpi = 300)


```

#Survival analysis for PTV burden group 4
```{r}

cancer_surv_PTV4 = lapply(X = unique(icd_mapping$cancer_type), FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group_20per = cut(temp_can$PTV_burden, breaks = quantile(temp_can$PTV_burden, c(0, 0.05, 0.95, 1)), labels = c("Low", "Medium","High"), include.lowest = T, right = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_4, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_4, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
})
ggarrange(plotlist =  cancer_surv_PTV4, ncol = 5, nrow = 5)

ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_burden_g4_surv_cancer.jpg", width = 100, height = 50, units = "cm", dpi = 300)

#temp_can$PTV_burden_group_4 = cut(temp_can$PTV_burden, breaks = quantile(temp_can$PTV_burden, c(0, 0.01, 0.99, 1)), labels = c("Low", "Medium","High"), include.lowest = T, right = T)

```

#All tumor survival analysis
```{r}
temp_df_1 = data.frame(eid = cancer_data$eid,
                       birthdate = cancer_data$birthdate,
                       date_of_death = cancer_data$date_of_death,
                       diag_date = cancer_data$diag_date,
                       death = cancer_data$death,
                       cancer_type = cancer_data$cancer_type,
                       PTV_burden_group = cancer_data$PTV_burden_group,
                       PTV_burden_group_2 = cancer_data$PTV_burden_group_2,
                       PTV_burden_group_3 = cancer_data$PTV_burden_group_3)

temp_df_1 %<>%
  group_by(eid) %>%
  summarize(
    birthdate = first(birthdate),
    date_of_death = first(date_of_death),
    diag_date = first(diag_date),
    death = first(death),
    PTV_burden_group = first(PTV_burden_group),
    PTV_burden_group_2 = first(PTV_burden_group_2),
    PTV_burden_group_3 = first(PTV_burden_group_3))

temp_df_1 %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
temp_df_1$event_indicator = ifelse(temp_df_1$death == T, 1, 0)

fit_4 = survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group, data = temp_df_1)
PTV_g1_cancer_surv_death = ggsurvplot(fit_4, risk.table = T, pval = T, tables.theme = clean_theme(), title = paste0("Survival Analysis for Tumors (PTV burden group 1)"))
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_g1_tumor_surv.png", plot = PTV_g1_cancer_surv_death$plot, width = 60, height = 30, units = "cm", dpi = 300)


fit_5 = survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_2, data = temp_df_1)
PTV_g2_cancer_surv_death = ggsurvplot(fit_5, risk.table = T, pval = T, tables.theme = clean_theme(), title = paste0("Survival Analysis for Tumors (PTV burden group 2)"))
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_g2_tumor_surv.png", plot = PTV_g2_cancer_surv_death$plot, width = 60, height = 30, units = "cm", dpi = 300)

fit_6 = survfit(Surv(time_in_days, event = event_indicator) ~ PTV_burden_group_3, data = temp_df_1)
PTV_g3_cancer_surv_death = ggsurvplot(fit_6, risk.table = T, pval = T, tables.theme = clean_theme(), title = paste0("Survival Analysis for Tumors (PTV burden group 3)"))
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_g3_tumor_surv.png", plot = PTV_g3_cancer_surv_death$plot, width = 60, height = 30, units = "cm", dpi = 300)



PTV_plot_list_surv = list(PTV_g1_cancer_surv_death$plot, PTV_g2_cancer_surv_death$plot, PTV_g3_cancer_surv_death$plot)
PTV_risk_list_surv = list(PTV_g1_cancer_surv_death$table, PTV_g2_cancer_surv_death$table, PTV_g3_cancer_surv_death$table)
PTV_plot_list_combined = c(PTV_plot_list_surv, PTV_risk_list_surv)

grid.arrange(grobs = PTV_plot_list_combined, ncol = 3)
plot_grid(plotlist = PTV_plot_list_combined, ncol = 3)

ggsave(filename ="C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_all_surv.png", width = 100, height = 50, units = "cm", dpi = 300 )


```

#PTV cox model
```{r}

coxregfun_surv = function(x) {
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden = factor(temp_can$PTV_burden, ordered = T)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden, data = temp_can)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value)
}

result_list_surv = lapply(unique(icd_mapping$cancer_type), coxregfun_surv)


cancer_new_PTV_surv_df = do.call(rbind, result_list_surv)
save(cancer_new_PTV_surv_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_surv")

```

#PTV cox model group 1
```{r}

coxregfun_surv_g1 = function(x) {
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group = factor(temp_can$PTV_burden_group, ordered = T)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden_group, data = temp_can)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value)
}

result_list_surv_g1 = lapply(unique(icd_mapping$cancer_type), coxregfun_surv_g1)


cancer_new_PTV_surv_g1_df = do.call(rbind, result_list_surv_g1)
save(cancer_new_PTV_surv_g1_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_surv_g1")

```

#PTV cox model group 2
```{r}

coxregfun_surv_g2 = function(x) {
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group_2 = factor(temp_can$PTV_burden_group_2, ordered = T)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden_group_2, data = temp_can)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value)
}

result_list_surv_g2 = lapply(unique(icd_mapping$cancer_type), coxregfun_surv_g2)


cancer_new_PTV_surv_g2_df = do.call(rbind, result_list_surv_g2)
save(cancer_new_PTV_surv_g2_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_surv_g2")

```

#PTV cox model group 3
```{r}

coxregfun_surv_g3 = function(x) {
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can$PTV_burden_group_3 = factor(temp_can$PTV_burden_group_3, ordered = T)
  res.cox = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden_group_3, data = temp_can)
  c(exp(res.cox$coefficients), summary(res.cox)$coefficients[5])
  p_value = summary(res.cox)$coefficients[5]
  coef_value = summary(res.cox)$coefficients[1]
  data.frame(
    cancer_type = x,
    p_value = p_value,
    coef_value = coef_value)
}

result_list_surv_g3 = lapply(unique(icd_mapping$cancer_type), coxregfun_surv_g3)


cancer_new_PTV_surv_g3_df = do.call(rbind, result_list_surv_g3)
save(cancer_new_PTV_surv_g3_df, file = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/objects/cancer_new_PTV_coef_surv_g3")

```

#Forset plot
```{r}

cancer_data$cancer_type[is.na(cancer_data$cancer_type)] = "No_cancer"
tumors = unique(cancer_data$cancer_type)
tumors = tumors[tumors != "No_cancer"]


i = 0
PTV_tumor_surv_forest_plot = lapply(X = tumors, FUN = function(x){
  i = i + 1
  .GlobalEnv$i = i
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can %<>% add_count(sex, name = "n_sex") %>% filter(n_sex >= 5)
  temp_can$sex = factor(temp_can$sex)
  l = temp_can %>% 
    add_count(ethnic_group, name = "freq") %>% 
    filter(freq > nrow(temp_can)/25) %>% 
    arrange(desc(freq)) %>% 
    pull(ethnic_group) %>% 
    unique()
  temp_can$ethnic_group = factor(temp_can$ethnic_group, levels = c("British", sort(setdiff(l, "British"))))
  temp_can %<>% 
    select(sex, ethnic_group, PTV_burden, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10, time_in_days,event_indicator) %>% 
    na.omit()
  if (length(l) == 1 & length(unique(temp_can$sex)) == 1){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  } else if(length(l) == 1 & length(unique(temp_can$sex)) == 2){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  } else if(length(l) > 1 & length(unique(temp_can$sex)) == 1){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + ethnic_group + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  } else {
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + ethnic_group + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  }
  ggforest(hr_model, data = as.data.frame(temp_can), main = paste0("Hazard ratio of ", x), refLabel = "reference", noDigits = 2)
})

PTV_surv_cox_forest = ggarrange(plotlist = PTV_tumor_surv_forest_plot, ncol = 9, nrow = 3)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/ptv_forest_surv.png", plot = PTV_surv_cox_forest, width = 150, height = 80, units = "cm", limitsize = F, dpi = 300)


```

#Forest plot hist
```{r}


cancer_data$hist_tumor_cat[is.na(cancer_data$hist_tumor_cat)] = "No_cancer"
hist_cat = unique(cancer_data$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

i = 0
PTV_tumor_surv_forest_plot = lapply(X = hist_cat, FUN = function(x){
  i = i + 1
  .GlobalEnv$i = i
  temp_can = subset(cancer_data, hist_tumor_cat == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$hist_tumor_cat[is.na(temp_can$hist_tumor_cat)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  temp_can %<>% add_count(sex, name = "n_sex") %>% filter(n_sex >= 5)
  temp_can$sex = factor(temp_can$sex)
  l = temp_can %>% 
    add_count(ethnic_group, name = "freq") %>% 
    filter(freq > nrow(temp_can)/25) %>% 
    arrange(desc(freq)) %>% 
    pull(ethnic_group) %>% 
    unique()
  temp_can$ethnic_group = factor(temp_can$ethnic_group, levels = c("British", sort(setdiff(l, "British"))))
  temp_can %<>% 
    select(sex, ethnic_group, PTV_burden, gpca1, gpca2, gpca3, gpca4, gpca5, gpca6, gpca7, gpca8, gpca9, gpca10, time_in_days,event_indicator) %>% 
    na.omit()
  if (length(l) == 1 & length(unique(temp_can$sex)) == 1){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  } else if(length(l) == 1 & length(unique(temp_can$sex)) == 2){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  } else if(length(l) > 1 & length(unique(temp_can$sex)) == 1){
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + ethnic_group + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  } else {
    hr_model = coxph(Surv(time_in_days, event_indicator) ~ PTV_burden + sex + ethnic_group + gpca1 + gpca2 + gpca3 + gpca4 + gpca5 + gpca6 + gpca7 + gpca8 + gpca9 + gpca10, data = temp_can)
  }
  ggforest(hr_model, data = as.data.frame(temp_can), main = paste0("Hazard ratio of ", x), refLabel = "reference", noDigits = 2)
})

PTV_surv_cox_forest = ggarrange(plotlist = PTV_tumor_surv_forest_plot, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/hist_ptv_forest_surv.png", plot = PTV_surv_cox_forest, width = 150, height = 80, units = "cm", limitsize = F, dpi = 300)

```

#PTV survival
##Universal 1
```{r}

cancer_data$hist_tumor_cat[is.na(cancer_data$hist_tumor_cat)] = "No_cancer"
hist_cat = unique(cancer_data$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

cancer_surv_PTV1 = lapply(X = hist_cat, FUN = function(x){
  temp_can = subset(cancer_data, hist_tumor_cat == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$hist_tumor_cat[is.na(temp_can$hist_tumor_cat)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_universal_1 = factor(temp_can$PTV_universal_1, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_universal_1, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_universal_1, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/hist_type_PTV_universal_1.png", width = 100, height = 50, units = "cm", dpi = 300)

```

##Universal 2
```{r}

cancer_data$hist_tumor_cat[is.na(cancer_data$hist_tumor_cat)] = "No_cancer"
hist_cat = unique(cancer_data$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

cancer_surv_PTV1 = lapply(X = hist_cat, FUN = function(x){
  temp_can = subset(cancer_data, hist_tumor_cat == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$hist_tumor_cat[is.na(temp_can$hist_tumor_cat)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_universal_2 = factor(temp_can$PTV_universal_2, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_universal_2, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_universal_2, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/hist_type_PTV_universal_2.png", width = 100, height = 50, units = "cm", dpi = 300)


```

##Hist type 1
```{r}

cancer_data$hist_tumor_cat[is.na(cancer_data$hist_tumor_cat)] = "No_cancer"
hist_cat = unique(cancer_data$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

cancer_surv_PTV1 = lapply(X = hist_cat, FUN = function(x){
  temp_can = subset(cancer_data, hist_tumor_cat == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can$PTV_hist_type_1 = as.character(temp_can$PTV_hist_type_1)
  l = sort(unique(temp_can$PTV_hist_type_1))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  temp_can$PTV_hist_type_1 = cut(temp_can$PTV_burden, breaks = cp, include.lowest = T, right = T)
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$hist_tumor_cat[is.na(temp_can$hist_tumor_cat)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_hist_type_1 = factor(temp_can$PTV_hist_type_1, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_1, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_1, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/hist_type_PTV_tumor_cat_1.png", width = 100, height = 50, units = "cm", dpi = 300)


```

##Hist type 2
```{r}
cancer_data$hist_tumor_cat[is.na(cancer_data$hist_tumor_cat)] = "No_cancer"
hist_cat = unique(cancer_data$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

cancer_surv_PTV1 = lapply(X = hist_cat, FUN = function(x){
  temp_can = subset(cancer_data, hist_tumor_cat == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can$PTV_hist_type_2 = as.character(temp_can$PTV_hist_type_2)
  l = sort(unique(temp_can$PTV_hist_type_2))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  temp_can$PTV_hist_type_2 = cut(temp_can$PTV_burden, breaks = cp, include.lowest = T, right = T)
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$hist_tumor_cat[is.na(temp_can$hist_tumor_cat)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_hist_type_2 = factor(temp_can$PTV_hist_type_2, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_2, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_2, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/hist_type_PTV_tumor_cat_2.png", width = 100, height = 50, units = "cm", dpi = 300)


```

##Hist type 3
```{r}
cancer_data$hist_tumor_cat[is.na(cancer_data$hist_tumor_cat)] = "No_cancer"
hist_cat = unique(cancer_data$hist_tumor_cat)
hist_cat = hist_cat[hist_cat != "No_cancer"]

cancer_surv_PTV1 = lapply(X = hist_cat, FUN = function(x){
  temp_can = subset(cancer_data, hist_tumor_cat == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can$PTV_hist_type_3 = as.character(temp_can$PTV_hist_type_3)
  l = sort(unique(temp_can$PTV_hist_type_3))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  temp_can$PTV_hist_type_3 = cut(temp_can$PTV_burden, breaks = cp, include.lowest = T, right = T)
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$hist_tumor_cat[is.na(temp_can$hist_tumor_cat)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_hist_type_3 = factor(temp_can$PTV_hist_type_3, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_3, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_hist_type_3, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 3, nrow = 2)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/hist_type_PTV_tumor_cat_3.png", width = 100, height = 50, units = "cm", dpi = 300)

```

##Cancer type 1
```{r}

cancer_data$cancer_type[is.na(cancer_data$cancer_type)] = "No_cancer"
tumors = unique(cancer_data$cancer_type)
tumors = tumors[tumors != "No_cancer"]

cancer_surv_PTV1 = lapply(X = tumors, FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can$PTV_cancer_type_1 = as.character(temp_can$PTV_cancer_type_1)
  l = sort(unique(temp_can$PTV_cancer_type_1))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  temp_can$PTV_cancer_type_1 = cut(temp_can$PTV_burden, breaks = cp, include.lowest = T, right = T)
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_cancer_type_1 = factor(temp_can$PTV_cancer_type_1, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_1, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_1, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/cancer_type_1_PTV.png", width = 100, height = 50, units = "cm", dpi = 300)

```

##Cancer type 2
```{r}

cancer_data$cancer_type[is.na(cancer_data$cancer_type)] = "No_cancer"
tumors = unique(cancer_data$cancer_type)
tumors = tumors[tumors != "No_cancer"]

cancer_surv_PTV1 = lapply(X = tumors, FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can$PTV_cancer_type_2 = as.character(temp_can$PTV_cancer_type_2)
  l = sort(unique(temp_can$PTV_cancer_type_2))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  temp_can$PTV_cancer_type_2 = cut(temp_can$PTV_burden, breaks = cp, include.lowest = T, right = T)
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_cancer_type_2 = factor(temp_can$PTV_cancer_type_2, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_2, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_2, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/cancer_type_2_PTV.png.png", width = 100, height = 50, units = "cm", dpi = 300)

```

##Cancer type 3
```{r}

cancer_data$cancer_type[is.na(cancer_data$cancer_type)] = "No_cancer"
tumors = unique(cancer_data$cancer_type)
tumors = tumors[tumors != "No_cancer"]

cancer_surv_PTV1 = lapply(X = tumors, FUN = function(x){
  temp_can = subset(cancer_data, cancer_type == x)
  temp_can = temp_can[!duplicated(temp_can$eid), ]
  temp_can$PTV_cancer_type_3 = as.character(temp_can$PTV_cancer_type_3)
  l = sort(unique(temp_can$PTV_cancer_type_3))
  cp = unique(sort(unlist(sapply(X = l, FUN = function(y)  sapply(X = strsplit(y, ",")[[1]], FUN = function(x) as.numeric(str_extract(x, "\\d+")))))))
  temp_can$PTV_cancer_type_3 = cut(temp_can$PTV_burden, breaks = cp, include.lowest = T, right = T)
  temp_can %<>% 
  mutate(time_in_days = case_when(
      death == T ~ date_of_death - diag_date,
      death == F ~ max(date_of_death, na.rm = T) - diag_date
    ))
  temp_can$cancer_type[is.na(temp_can$cancer_type)] = "No cancer"
  temp_can$event_indicator = ifelse(temp_can$death == T, 1, 0)
  #temp_can$PTV_cancer_type_3 = factor(temp_can$PTV_cancer_type_3, ordered = T)
  fit = do.call(survfit, list(formula = Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_3, data = temp_can))
  args = list(fit = fit, title = paste0("Survival Analysis for ", x, "\n(N=", nrow(temp_can), ")"), pval = T, legend = "none")
  pfit = do.call(ggsurvplot, args)
  tblA = ggsurvplot(survfit(Surv(time_in_days, event = event_indicator) ~ PTV_cancer_type_3, data = temp_can), risk.table = T, tables.theme = clean_theme())$table
  grid.arrange(pfit$plot, tblA, layout_matrix = rbind(c(1, 1), c(1,1), c(2, 2)))
  })


ggarrange(plotlist =  cancer_surv_PTV1, ncol = 5, nrow = 5)
ggsave(filename = "C:/Users/bagil/Desktop/MyGit/ukb_tumor/plots/Survival/PTV_cancer_surv/cancer_type_3_PTV.png.png", width = 100, height = 50, units = "cm", dpi = 300)

```

